### 2024-01-04 conference/IGARSS2024.qmd
update for igarrs2023

---
title: "DDS4Chl: A dataset of drought indices of demand and water supply for continental Chile"
format:
  ieee-pdf:
    keep-tex: false  
  ieee-html: default
title-block-banner: true
author:
  - id: dfolio
    name: Francisco Zambrano
    affiliations:
      - name: Facultad de Ciencias, Universidad Mayor
        department: Hemera Centro de Observación de la Tierra
        city: Santiago
        country: Chile
        postal-code: 8580745
        orcid: 0000-0001-6896-8534
        email: francisco.zambrano@umayor.cl
        #url: https://dfolio.fr/
        attributes:
        corresponding: true
        #photo: david-folio.png
  - id: iduran
    name: Iongel Duran
    affiliations: 
      - name: Facultad de Ciencias, Universidad Mayor
        department: Hemera Centro de Observación de la Tierra
        city: Santiago
        country: Chile
        postal-code: 8580745

note: "Template created June 23, 2023; revised `r format(Sys.Date(),format='%B %d, %Y')`."
#keywords: [IEEE, IEEEtran, journal, Quarto, Pandoc, template]
#funding: 
funding: 
  statement: "The `quarto-ieee` template is freely available under the MIT license on github: <https://github.com/dfolio/quarto-ieee>."
pageheader:
  left: IGARSS, January 2024
  right: 'D. Folio:  A Sample Article Using quarto-ieee'
#bibliography: bibliography.bib
date: 2024-01-05
pdf: https://github.com/dfolio/quarto-ieee/blob/main/template.pdf
citation: 
  container-title: GitHUB
  page: 1-3
  type: software
  issued: 2023-06-23
  url: https://github.com/dfolio/quarto-ieee
  pdf-url: https://github.com/dfolio/quarto-ieee/template.pdf
---

# **Abstract** {-}

Central Chile has been affected by a persistent drought which is impacting the hydrological system and vegetation development. The region has been the focus of research studies due to a persistent period of water scarcity which has been defined as a “mega drought.  
<br>

**Index Terms—** One, two, three, four, five

# **Introduction**

# **Methods**

# **Results**

# Conlusion