### 2024-03-12 article/paper_drought_Chile2023.qmd
update paper qm intro

---
title: Drought, vegetation productivity, and land cover change in Chile
#subtitle: A Short Subtitle
author:
  - name: Francisco Zambrano
    email: francisco.zambrano@umayor.cl
    affiliations: 
        - id: umayor
          name: Universidad Mayor
          department: Hémera Centro de Observación de la Tierra, Facultad de Ciencias, Escuela de Ingeniería en Medio Ambiente y Sustentabilidad
          #address: Street Address
          city: Santiago, Chile
          #state: State
          postal-code: 7500994
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Anton Vrieling
    email: a.vrieling@utwente.nl
    affiliations:
        - id: itc
          name: University of Twente
          department: Faculty of Geo-Information Science and Earth Observation (ITC)
          #address: Street Address
          city:  Enschede, The Netherlands
  - name: Francisco Fernández
    email: a.vrieling@utwente.nl
    affiliations:
        - id: uss
          name: Universidad San Sebastian
          department: 
          #address: Street Address
          city:  
  - name: Francisco Meza
  - name: Alejandro Venegas-González
  - name: Iongel Duran-Llacer
  - name: Nicolas Raab
  - name: Dylan Craven        
  #   note: |
  #     Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
  # - name: Cat Memes
  #   email: cat@example.com
  #   affiliations:
  #       - ref: another-u
  #   note: Yet another author footnote.
  # - name: Derek Zoolander
  #   email: derek@example.com
  #   affilations:
  #       - ref: some-tech
abstract: |
  Central Chile has been the focus of research studies due to the persistent decrease in water supply, which is impacting the hydrological system and vegetation development. This persistent period of water scarcity has been defined as a megadrought. Our objective is to assess the impact of drought on LULCC (land use land cover change) over continental Chile using drought indices of water supply and demand, soil moisture, and their impact on vegetation productivity. The monthly ERA5-Land (ERA5L) variables for precipitation, temperature, and soil moisture were used. From 2001 to 2022, we used the land cover MODIS product MCD12Q1, and from 2000 to 2023, we used the NDVI (Normalized Difference Vegetation Index) product MOD13A3 collection 6.1. As drought indices, we compute the standardized anomaly of cumulative NDVI (zcNDVI), the Standardized Precipitation Evapotranspiration Index (SPEI), the Evaporative Demand Drought Index (EDDI), and the Standardized Soil Moisture Index (SSI). These indices were calculated for time scales of 1, 3, 6, 12, 24, and 36 months, except for zcNDVI, which was for 6 months. We analyze the trend for LULCC, vegetation productivity, and drought indices. Also, we analyzed the temporal correlation of SPI, SPEI, EDDI, and SSI with zcNDVI to gain insights into the impact of water supply and demand on vegetation productivity. Our results showed that LULCC was highest in "Centro," "Sur," and "Austral,” with 36%, 31%, and 34% of change in the surface type, respectively. The EDDI shows that water demand has increased for all zones, with a major increase in "Norte Grande." The drought indices of water supply and soil moisture evidence a decreasing trend, which decreases at longer time scales, from "Norte Grande" to "Sur." "Austral" is the only zone that shows an increase in supply. Vegetation productivity measures by zcNDVI present a negative trend in "Norte Chico" and "Centro." On the other hand, forests seem to be the most resistant to drought. The types that show to be most affected by variation in climate conditions are shrublands, savannas, and croplands. The drought indices that have the capability of explaining to a major degree the variance in vegetation productivity are SSI-12, followed by SPEI-24 and SPEI-12 in "Norte Chico" and "Centro." The results indicate that “Norte Chico” and “Zona Central” are the most sensitive regions to water supply deficits lasting longer than a year. Our results can help develop a robust vegetation productivity forecasting model for land cover classes in Chile.
keywords: 
  - drought
  - land cover change
  - satellite
date: last-modified
bibliography: references.bib
keep-tex: false
include-before-body:
  - text: |
      \captionsetup{justification=raggedright,singlelinecheck=false}
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      layout: onecolumn
      model: 3p
      cite-style: authoryear
      natbiboptions: numbers
    include-in-header: 
      text: \usepackage{lineno}\linenumbers
            \usepackage{multirow}
            \usepackage{lscape}
            \newcommand{\blandscape}{\begin{landscape}}
            \newcommand{\elandscape}{\end{landscape}}
  docx:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      layout: onecolumn
      model: 3p
      cite-style: authoryear
      natbiboptions: numbers
    include-in-header: 
      text: \usepackage{lineno}\linenumbers
            \usepackage{multirow}
            \usepackage{lscape}
            \newcommand{\blandscape}{\begin{landscape}}
            \newcommand{\elandscape}{\end{landscape}}
---

# Introduction

Drought is often classified as meteorological when there is a decrease in precipitation below the mean average of several years (more than 30 years), hydrological when these anomalies last for long periods (months to years) and affect water systems, and agricultural when the deficit impacts plant health anomalies and leads to decreased productivity [@Wilhite1985]. However, it is important to note that drought is also influenced by human activities, which were not considered in the definitions. Thus, @Loon2016 and @AghaKouchak2021 have given an updated definition of drought for the Anthropocene, suggesting that it should be considered the feedback of humans’ decisions and activities that drives the anthropogenic drought. Simultaneously, drought leads to heightened tree mortality and induces alterations in land cover and land use, ultimately affecting ecosystems [@Crausbay2017]. Even though many ecological studies have misinterpreted how to characterize drought, for example, sometimes considering "dry" conditions as "drought" [@Slette2019]. Then, @Crausbay2017 proposed the ecological drought definition as "an episodic deficit in water availability that drives ecosystems beyond thresholds of vulnerability, impacts ecosystem services, and triggers feedback in natural and/or human systems." In light of current global warming, it is crucial to study the interaction between drought and ecosystems in order to understand their feedback and impact on water security. [@Bakker2012]

Human-induced greenhouse gas emissions have increased the frequency and/or intensity of drought as a result of global warming, according to the sixth assessment report (AR6) of the Intergovernmental Panel on Climate Change (IPCC) [@IPCC2023]. The evidence supporting this claim has been strengthened since AR5 [@IPCC2013]. Recent studies, however, have produced contrasting findings, suggesting that drought has not exhibited a significant trend over the past forty years. [@Vicente-Serrano2022; @Kogan2020]. @Vicente-Serrano2022 analyzed the meteorological drought trend on a global scale, finding that only in a few regions has there been an increase in the severity of drought. Moreover, they attribute the increase in droughts over the past forty years solely to an increase in atmospheric evaporative demand (AED), which in turn enhances vegetation water demand, with important implications for agricultural and ecological droughts. Also, they state that "the increase in hydrological droughts has been primarily observed in regions with high water demand and land cover change". Similarly, @Kogan2020 analyzed the drought trend using vegetation health methods, finding that for the globe, hemispheres, and main grain-producing countries, drought has not expanded or intensified for the last 38 years. Further, the @IPCC2021 suggests that there is a high degree of confidence that rising temperatures will increase the extent, frequency, and severity of droughts. Also, AR6 [@IPCC2023] predicts that many regions of the world will experience more severe agricultural and ecological droughts even if global warming stabilizes at 1.5°–2°C. To better evaluate the impact of drought trends on ecosystems, assessments are needed that relate meteorological and soil moisture variables to their effects on vegetation.

From 1960 to 2019, land use change has impacted around one-third of the Earth's surface, which is four times more than previously thought [@Winkler2021]. Multiple studies aim to analyze and forecast changes in land cover globally [@Winkler2021; @Song2018] and regionally [@Chamling2020; @Homer2020; @Yang2021]. Some others seek to analyze the impact of land cover change on climate conditions such as temperature and precipitation [@Luyssaert2014;@Pitman2012]. There is less research on the interaction between drought and land cover change [@Chen2022;@Akinyemi2021;@Peng2017]. @Peng2017 conducted a worldwide investigation utilizing net primary production to examine the spatial and temporal variations in vegetation productivity at global level. The study aimed to assess the influence of drought by comparing the twelve-month Standardized Precipitation Evapotranspiration Index (SPEI) and land cover change. According to their findings, drought is responsible for 37% of the decline in vegetation productivity, while water availability accounts for 55% of the variation. @Chen2022 studied the trend of vegetation greenness and productivity and its relation to meteorological drought (SPEI of twelve months in December) and soil moisture at the global level. The results showed lower correlations (<0.2) for both variables. @Akinyemi2021 evaluates drought trends and land cover change using vegetation indices in Botwsana in a semi-arid climate. These studies mostly looked at how changes in land cover and vegetation productivity are related to a single drought index (SPEI) over a single time period of 12 months. SPEI takes into account the combined effect of precipitation and AED as a water balance, but it does not allow us to know the contribution of each variable on its own. Some things worth investigating in terms of land cover change and vegetation productivity are: i) How do they respond to short- to long-term meteorological droughts? ii) How do they behave in humid and arid climatic zones regarding drought? And iii) What is the role of soil moisture? Likewise, there is a lack of understanding of how the alteration in water supply and demand is affecting land cover transformations.

Chile's diverse climatic and ecosystem types (@Beck2023;@Luebert2022) make it an ideal natural laboratory for studying climate and ecosystems. Additionally, the country has experienced severe drought conditions that have had significant effects on vegetation and water storage. Central Chile faced a persistent precipitation deficit between 2010 and 2022, defined as a megadrought [@Garreaud2017], which has impacted the Chilean ecosystem. This megadrought was defined by the Standardized Precipitation Index (SPI) of twelve months in December having values below one standard deviation. Some studies have addressed how this drought affects single ecosystems in terms of forest development [@Miranda2020;@Venegas2018], forest fire occurrence [@UrrutiaJalabert2018], and crop productivity [@Zambrano2023;@Zambrano2018;@Zambrano2016]. We found one study regarding land cover and drought in Chile. The study by @Fuentes2021  evaluates water scarcity and land cover change in Chile between 29° and 39° of south latitude. @Fuentes2021 used the SPEI of one month for evaluating drought, which led to misleading results. For example, they did not find a temporal trend in the SPEI but found a decreasing trend in water availability and an increase trend on AED, which in turn should have been capable of being captured with longer time scales of the SPEI. The term "megadrought" in Chile is used to describe a prolonged water shortage that lasts for several years, resulting in a permanent deficit that impacts the hydrological system [@Boisier2018]. Hence, it is imperative to assess temporal scales that take into account the cumnulative effect within some years. There is little knowledge about the relationship between drought and ecosystem in Chile; thus, it is important to understand in more detail how meteorological and soil moisture droughts influence ecosystem dynamics to inform adaptation options.

A detailed spatiotemporal assessment of the interaction of drought for short- to long-term and land cover change requires information on vegetation as well as weather variables such as precipitation, temperature, and soil moisture. Weather networks in Chile present some disadvantages, such as spatio-temporal gaps, a short history, and irregular quality, which make them difficult to represent the whole extent of the country spatially. In order to do this, we use reanalysis data from ERA5-Land [@MunozSabater2021] to create drought indices that consider AED, precipitation, and soil moisture over a range of time periods, from the short to the long term. Also, we use vegetation spectral information and annual land cover change from the Moderate-Resolution Imaging Spectroradiometer (MODIS). We expect to gain insight regarding the temporal evolution of water demand, water supply, and soil moisture, as well as the interaction with land cover change and vegetation productivity. Here, we analyze the multi-dimensional impacts of drought across ecosystems in continental Chile. More specifically, we aim to assess: i) temporal changes in land-use cover and the direction and magnitude of their relationships with drought indices for water demand and supply, soil moisture, and vegetation productivity; ii) short- to long-term temporal trends in multi-scalar drought indices; and iii) the relationship between vegetation productivity and drought indices for water demand and supply and soil moisture across Chilean ecosystems.

# Study area

Continental Chile has a diverse climate conditions with strong gradients from north to south and east to west [@Aceituno2021] (@fig-studyArea a), which determines its great ecosystem diversity (@Luebert2022) (@fig-studyArea c). The Andes Mountains are a main factor in climate latitudinal variation [@Garreaud2009]. In order to characterize the climate and ecosystem of Chile, we utilize the Köppen-Geiger classification system developed by @Beck2023 and the land cover data derived from the MODIS product for the period of 2001–2022, based on the International Geosphere-Biosphere Programme (IGBP) classification scheme proposed by @Friedl2019. "Norte Grande" and "Norte Chico" predominate in an arid desert climate with hot (Bwh) and cold (Bwk) temperatures. At the south of "Norte Chico," the climate changes to an arid steppe with cold temperatures (Bsk). In these two northern regions, the land is mostly bare, with a minor surface of vegetation types such as shrubland and grassland. In the zones "Centro" and the north half of "Sur," the main climate is Mediterranean, with warm to hot summers (Csa and Csb). Land cover in "Centro" comprises a significant amount of shrubland and savanna (50%), grassland (16%), forest (8%), and croplands (5%). An oceanic climate (Cfb) predominates in the south of "Sur" and the north of "Austral." Those zones are high in forest and grassland. The southern part of the country has a tundra climate, and in "Austral", it is a cold semi-arid area with an extended surface of grassland, forest, and, to a lesser extent, savanna.

```{r fig-studyArea}
#| echo: false
#| fig-pos: "!ht"
#| fig-env: 'figure*'
#| fig-cap: '(a) Chile with the Koppen-Geiger climate classes and the five macrozones "Norte Grande", "Norte Chico",  "Centro", "Sur", and "Austral". (b) Topography reference map. (c) land cover classes for 2022. (d) Persistent land cover classes (> 80%) for 2001-2022'
#| out-width: "\\textwidth"
knitr::include_graphics("../output/figs/map_study_con_landcover.png")
```

# Materials and Methods

## Software and packages used

For the downloading, processing, and analysis of the spatio-temporal data, we used the open source software for statistical computing and graphics, `R` [@R2023]. For downloading ERA5L, we used the `{ecmwfr}` package [@Hufkens2019]. For processing raster data, we used `{terra}` [@Hijmans2023] and `{stars}` [@Pebesma2023]. For managing vectorial data, we used `{sf}` [@Pebesma2018]. For the calculation of AED, we used `{SPEI}` [@Bergueria2023]. For mapping, we use {tmap} [@Tennekes2018]. For data analysis, the suite {tidyverse} [@Wickham2019] was used.

##	Data

###	Earth observation data

For water supply and demand variables, we used ERA5L [@MunozSabater2021], a reanalys dataset that provides the evolution of land variables since 1950. It has a spatial resolution of 0.1°, hourly frequency, and global coverage. We selected the variables for total precipitation, 2 meter temperature maximum and minimum, and volumetric soil water layers between 0 and 100cm of depth (layer 1 to layer 3). The data was downloaded using the Copernicus Climate Data Store (CDS) Application Program Interface (API) implemented in `{ecmfwr}` [@Hufkens2019].

To derive a proxy for vegetation productivity, we used the product MOD13A3 collection 6.1 from MODIS [@Didan2015]. It provides vegetation indices (NDVI and EVI) at 1km of spatial resolution and monthly frequency. The MOD13A3.061 and MCD12Q1.061 were retrieved from the online Data Pool, courtesy of the NASA EOSDIS Land Processes Distributed Active Archive Center (LP DAAC), USGS Earth Resources Observation and Science (EROS) Center, Sioux Falls, South Dakota, https://lpdaa.usgs.gov/tools/data-pool/.

```{=latex}
\begin{table}[!ht]
\caption{Description of the earth observation data used }
\label{tab-desEOD}
\small
\centering
\begin{tabular}{p{0.13\textwidth}cp{0.3\textwidth}p{0.095\textwidth}ccc}
\hline
\multirow{1}{*}{\centering Product} & Sub-product & Variable & Spatial Resolution  & Period & Units & Short Name \\ 
\hline
\multirow{4}{*}{ERA5L} & ~ & Precipitation & \multirow{4}{*}{~0.1°} & \multirow{4}{*}{1981-2023} & mm & P \\ 
         &  & Maximum temperature & ~ & & $°C$ & $T_{max}$ \\ 
         &  & Minimum temperature & ~ & & $°C$ & $T_{min}$ \\ 
         &  & Volumetric Soil Water Content at 1m & ~ & & $m3/m3$ & SM \\ 
ERA5L* & & Atmospheric Evaporative Demand & 0.1° & 1981-2023 & mm & AED \\
        \multirow{2}{*}{MODIS} & MOD13A3.061 & Normalized Difference Vegetation Index & \multirow{2}{*}{~1 km} & 2000-2023 & ~ & NDVI \\ 
         & MCD12Q1.061 & land cover IGBP scheme & & 2001-2022 & ~ & land cover \\ 
\hline
\end{tabular}
{\raggedright *Derived from ERA5L with Eq. \ref{eq-AED}. \par}
\end{table}
```

###	Weather stations

We compared the ERA5L variables for monthly mean temperature, total precipitation, and volumetric soil water content against values retrieved by weather stations. For temperature and precipitation, we used the weather network from the Ministry of Agriculture of Chile (www.agromet.com) between 2015 and 2023. We used 277 stations located throughout Chile. For soil moisture, we select a private soil network that is owned by the agricultural enterprise Garces Fruit, which has 99 stations in Central Chile, located in cherry fruit crops. The sensors are installed at 30, 60, and 90m and are the model Teros 12 from MeterGroup. To avoid the effect of irrigation on soil moisture, which ERA5L hardly captures, we used daily data for the year 2022 and the months outside the growing season, May to September.

### Validation of ERA5L variables

To account for the performance of the ERA5L climatic variables regarding the values measured by the weather stations. We selected the following metrics:

$$MAE = \frac{1}{n}\sum |{E-S}|$$
$$Bias = \frac{\sum E}{\sum S}$$
$$ubRMSE =\sqrt{\frac{\sum{ \left[ (E_i-\overline{E})-(S_i-\overline{S}) \right ] ^2}}{n}}$$

$$CC = \frac{\sum (S_i-\bar{S})(E_i-\bar{E})}{\sqrt{(Si-\bar{S})^2(E_i-\bar{E})^2}}$$

$MAE$: mean absolute error
$bias$: bias
$ubRMSE$: unbiassed root mean squared error
$CC$: coefficient of correlation
$S$: value of the variable measure by the weather station
$E$: value of the variable measure by ERA5L

## Drought Indices

### Atmospheric Evaporative Demand (AED)

For the indices EDDI and SPEI that use water demand, first we have to calculate the AED. For this, we used the method of Hargreaves [@Hargreaves1994;@Hargreaves1985]:

$$AED = 0.0023\cdot Ra\cdot (T+17.8)\cdot (T_{max}-T_{min})^{0.5}$$ {#eq-AED}

where $Ra$ $(MJ\,m^2\, day^{-1})$ is extraterrestrial radiation; $T$, $T_{max}$, and $T_{min}$ are mean, maximum, and minimum temperature $(°C)$. We calculate the centroid coordinates per pixel and use the latitude to estimate $Ra$.

We chose the method of Hargreaves to estimate AED because of its simplicity, which only requires temperatures and extrarrestrial radiation. Also, it has been recommended over other methods when the use of several climatic variables is limited [@Vicente-Serrano2014].

### Non-parametric calculation of drought indices

We derived the drought indices of water supply and demand, soil moisture from the ERA5L dataset, and vegetation from the MODIS product, all at monthly frequency.

To evaluate water demand, we chose the $EDDI$ [@Hobbins2016; @McEvoy2016] index, which uses the $AED$. We used the SPI [@Mckee1993], a drought monitoring index that the World Meteorological Organization (WMO) recommends. We calculated the SPEI, which used a balance between $P$ and $AED$, in this case, an auxiliary variable $D = P-AED$ is used. In this study, we used the $SSI$ (standardized soil moisture index at 1 m) [@Hao2013;@AghaKouchak2014], which uses soil moisture at 1m depth. Finally, for the proxy of productivity, $zcNDVI$, we used the NDVI. Before using the NDVI, it was smoothed using a locally weighted polynomial regression, following the procedure described in @Zambrano2018 and @Zambrano2016.

All the indices are multi-scalar and were calculated for time scales of 1, 3, 6, 12, 24, and 36 months, except for zcNDVI, which was calculated for 6 months. The goal is to be able to evaluate short- and long-term droughts in water demand and supply and soil moisture. This is particularly important for central Chile because it has suffered from a prolonged decrease in precipitation for more than 12 years [@Garreaud2020;@Boisier2018;@Garreaud2017].

To calculate the drought indices, first we must calculate the accumulation of the variable. In this case, for generalization purposes, we will use $V$, referring to $P$, $AED$, $D$, $NDVI$, and $SM$ (Table \ref{tab-desEOD}). We cumulated each $V$ over the time series of $n$ values, and for the time scales $s$:

$$A_{si} = \sum_{i=n-s-i+2}^{n-i+1} V_i\,\, \forall\, i\geq n-s+1  $$ {#eq-sumvar}

It corresponds to a moving window (convolution) that sums the variable for $s$ starting for the last month $n$ until the month, which could sum for $s$ months (n-s+1). Once the variable is cumulated over time for the scale $s$, we used a nonparametric approach following @Hobbins2016 to derive the drought indices. Thus, the empirically derived probabilities are obtained through an inverse normal approximation [@Abramowitz1968]. Then, we used the empirical Tukey plotting position [@Wilks2011] over $A_i$ to derive the $P(A_i)$ probabilities across a period of interest:

$$P(A_i) = \frac{i-0.33}{n+0.33'}$$ {#eq-probPai}

The drought indices $SPI$, $SPEI$, $EDDI$, $SSI$, and $zcNDVI$ are obtained following the inverse normal approximation:

$$DI(A_i) = W - \frac{C_0+C_1\cdot W + c_2 \cdot W^2}{1+d_1\cdot W +d_2\cdot W^2 +d_3\cdot W^3}$$ {#eq-DI}

$DI$ is referring to the drought index calculated for the variable $V$. The values for the constats are: $C_0 = 2.515517$, $C_1 = 0.802853$, $C_2 = 0.010328$, $d_1 = 1.432788$, $d_2 = 0.189269$, and $d3 = 0.001308$. For $P(A) \leq 0.5$, W=$\sqrt{-2\cdot ln(P(A_i))}$ , and for $P(A_i) > 0.5$, replace $P(A_i)$ with $1-P(A_i)$ and reverse the sign of $DI(A_i)$.

##	LULC change for 2001-2022 and its relation with water supply and demand, and soil moisture {#sec-methods_lulc}

### Land cover macroclasess and validation

To analyze the LULCC, we use the IGBP scheme from the MCD12Q1 collection 6.1 from MODIS. This product has a yearly frequency from 2001 to 2022. The IGBP defines 17 classes; from these, we regrouped into ten macroclasses, as follows: classes 1-4 to forest, 5-7 to schrublands, 8-9 to savannas, 10 as grasslands, 11 as wetlands, 12 and 14 to croplands, 13 as urban, 15 as snow and ice, 16 as barren, and 17 to water bodies. Thus, we have a land cover raster time series with the ten classes for 2001 and 2023.

To validate the land cover obtained, we compare the macroclasses with the ones of a more detailed land cover map made by @Zhao2016 for Chile with samples acquired in the years 2013–2014 (LCChile). The later has a spatial resolution of 30 m and three levels of defined classes; from those, we used level 1, which fits with the macroclasses land cover. We chose the years 2013 (IGBP2013) and 2014 (IGBP2014) from land cover macrolcasses to validate with LCChile.

We follow the next procedure:

i) resampled LCChile to the spatial resolution (500m) of the land cover macroclasses using the nearest neighbor method,
ii) took a random sample of 1000 points within continental Chile and extracted the classes that fell within each point for LCChile, IGBP2013, and IGBP2014; we considered the point extracted from LCChile as the truth and the values from the other two years as predictions.
iii) calculate a confusion matrix with the classes extracted from the 1000 points for LCChile, IGBP2013, and IGBP2014. Calculate the performance metrics of accuracy and F1.

$$Accuracy = \frac{TP+TN}{TP+TN+FP+FN}=\frac{correct\,\, classifications}{all\,\, classifications}$$
$$F1=\frac{2\cdot TP}{2\cdot TP + FP +FN}$$
where $TP$ and $FN$ refer to true positive and false negative, correctly classified classes; $TN$ and $FP$ to true negative and false positive, wrongly classified classes.

### Land cover persistence mask 2001-2022

Climate, vegetation development, seasonality, and changes in vegetation type all have an impact on the time series of NDVI. In this study, we want to examine the variation in vegetation productivity across various land cover types and how water demand, water supply, and soil moisture affect it. In order to avoid changes due to a change in the land cover type that will wrongly impact NDVI, we will develop a persistence mask for land cover for 2001–2023. Thereby, we reduce an important source of variation on a regional scale.

Therefore, we generated a raster mask for IGBP MODIS per pixel using macroclasses that remain constant for at least 80% of the years (2001–2022). This enables us to identify regions where the land cover of the macroclasses has remained constant.

### Land cover trend and drought indices

We calculated the surface occupied per land cover class into the five macrozones ("Norte Grande" to "Austral") per year for 2001–2023. After that, we calculated the trend's change in surface; we used the Sen' slope [@Sen1968] based on Mann-Kendall [@Kendall1975]. This way, we obtain a matrix of trends of 5 x 5 (macrozones x land cover). The aim is to later explore if the trend in land cover classes is associated with a trend in the drought indices. For this, we will use the techniques of regresion and regularization of Lasso [@Tibshirani2010] and Ridge [@Hoerl1970]. Also, we will test random forests for this purpose [@Ho1995]. We will choose the trend of land cover surface per macroclass and macrozone as the response variable and the trend of the drought indices (SPI, SPEI, EDDI, and SSI for time scales 1, 3, 6, 12, 24, and 36 months) as the predictor variables. With this analysis, we expect to gather insights regarding whether there is a pattern of climatic influence along Chile or if what is happening in Central Chile has to do with more localized climatic conditions.

## Trend of drought indices for water demand and supply, soil moisture, and vegetation productivity

### Mann-Kendall and Sen's slope

To estimate if there are significant positive or negative trends for the drought indices, we used the non-parametric test of Mann-Kendall [@Kendall1975]. To determine the magnitude of the trend, we used Sen's slope [@Sen1968]. Some of the advantages of applying this methodology are that the Sen's slope is not affected by outliers as regular regression does, and it is a non-paramteric method that is not affected by the distribution of the data. We applied both to the six time scales from 1981 to 2023 (monthly frequency) and the indices SPI, EDDI, SPEI, and SSI. In the case of zcNDVI (six months) was for 2000 to 2023. Thus, we have 31 trends. Also, we extracted the trend aggregated by macrozone and land cover class, obtaining a table of 31x5x5 (drought indices trends x macrozone x land cover class). We will use this data in @sec-methods_lulc to analyze if there is a strong relationship between the trends of drought indices and land cover surface within continental Chile.

### Trend in vegetation productivity without land cover change

@Vicente-Serrano2021 made a global analysis of the drought's severity trend using SPI, SPEI, and the Standardized Evapotranspiration Deficit Index (SEDI; @Vicente-Serrano2018) to evaluate AED. They indicate that the increase in hydrological drought has been due to anthropogenic effects rather than climate change. This is because the global increase in AED did not explain the change in the spatial pattern of the hydrological drought. Also, they state that _"the increase in hydrological droughts has been primarily observed in regions with high water demand and land cover change"_. We will contrast this hypothesis with what is occurring in Chile. To achieve this, we will use the land cover class type that remains more than 80% of types for 2001–2022 to evaluate the trend on zcNDVI and use this as a mask where there are low changes.

##	Impact for water supply and demand, and soil moisture in vegetation productivity within land cover types

We analyze the drought indices of water demand and supply and soil moisture against vegetation to address: i) if short- or long-term time scales are most important in impacting vegetation through Chile; and ii) the strength of the correlation for the variable and the time scale. Then, we will summarize for each land cover class and macrozone. Thus, we will be able to advance in understanding how climate is affecting vegetation, considering the impact on the five macroclasses of vegetation: forest, cropland, grassland, savanna, and shrubland.

An analysis is conducted on the linear correlation between the indices SPI, SPEI, EDDI, and SSI over time periods of 1, 3, 6, 12, 24, and 36 months, and zcNDVI. The objective is to determine the impact of soil moisture and water demand and supply on vegetation productivity. We implemented a methodology similar to that of @Meroni2016 when comparing the SPI for meteorological drought to the cumulative FAPAR (Fraction of Absorbed Photosynthetically Active Radiation), which served as an indicator for vegetation productivity. A pixel-to-pixel linear correlation analysis was performed for each index. To begin, the Pearson coefficient of correlation is computed for each of the six time scales. A significant time scale is identified as the one that attains the highest correlation (p < 0.05). Subsequently, the Pearson correlation coefficient corresponding to the time scales at which the value peaked was extracted. As a result, for each index, we generated two raster maps: one containing the time scales and the other the correlation value.

#	Results

## Data

### Validation of ERA5L variables

The average metrics of performance of ERA5L over the 266 weather stations were in the case of monthly temperature: $ubRMSE=1.06\,°C$, $MAE=1.131\,°C$, and $CC=0.963$, showing a good agreement, low error, and low overestimation. For cumulative monthly precipitation, $MAE=28.1\,mm$, $bias=1.93$, and $CC=0.845$, showing a high correlation and a 93% bias and being overestimated by ERA5L. In the case of the 97 soil moisture stations, we averaged for the three depths (30, 60, and 90m) and then compared it with volumetric water content at 1m derived from ERA5L. For this case, we made a daily comparison, having a $CC=0.71$, $RMSE=0.174\,m^3m^{-3}$, $MAE=0.167\,m^3m^{-3}$, and $bias=1.74$. The ERA5 soil moisture overestimate is 74%, but it has a kind of good correlation. For more detailed information, consult the supplementary material (SSX).

##	LULC change for 2001-2022 and its relation with water supply and demand, and soil moisture

### Land cover macroclasess and validation

For vegetation, we obtained and use hereafter five macroclasses of land cover from IGBP MODIS: forest, shrubland, savanna, grassland, and croplands. @fig-studyArea c shows the spatial distribution of the macroclasses through Chile for the year 2022. The validation of IGBP2013 and IGBP2014 with LCChile reached near the same metrics of performance, having an accuracy of ~0.82 and a F1 score of ~0.66 (see SS1).

### Land cover persistence mask 2001-2022 {#sec-persistence_mask}

@fig-studyArea d, shows the macroclasses of land cover persistance (80%) during 2021-2022, respectively (\ref{tab-landcoverSurf}). Within continental Chile, forest is the vegetation type with the highest surface area at 135,00 $km^2$, followed by grassland (73,176 $km^2$), savanna (54,410 $km^2$), shrubland (24,959 $km^2$), and cropland (3,100 $km^2$) (Table \ref{tab-landcoverSurf}). The macrozones with major LULCC for 2001–2022 were "Centro," "Sur," and "Austral" with 36%, 31%, and 34%, respectively (@fig-studyArea and Table \ref{tab-landcoverTrend}); of its surface that changes the type of land cover. @fig-LCprop shows the summary of the proportion of surface per land cover class and macrozone, derived from the persistance mask over continental Chile.

```{=latex}
\begin{table}[!ht]
\caption{Surface of the land cover class that persist during 2001-2022}
\label{tab-landcoverSurf}
\includegraphics[width = .5\textwidth]{../output/figs/table_surface_landcover_macrozone.png}
\end{table}
```

```{r fig-LCprop}
#| echo: false
#| fig-pos: "!ht"
#| fig-cap: "Proportion of land cover class from the persistent land cover for 2001-2022 (>80%) per macrozone"
#| out-width: "\\textwidth"
knitr::include_graphics("../output/figs/LC_pers80_per_macrozone.png")
```

### Land cover trend and drought indices

\begin{table}[!ht]
\caption{The value of Sen's slope trend next to the time-series plot of surface per land cover class (IGBP MCD12Q1.016) for 2001–2022 through Central Chile. Values of zero indicate that there was not a significant trend. Red dots on the plots indicate the maximum and minimum values of surface.}
\label{tab-landcoverTrend}
\includegraphics[]{../output/figs/table_var_landcover_macro.png}
\end{table}

The "Norte Chico" shows an increase in barrend land of 111 $km^2 year^{-1}$ and a reduction in the class savanna of 70 $km^2 year^{-1}$. In the "Centro" and "Sur," there are changes in the Chilean matorral, with an important reduction in savanna (136 to 318 $km^2 yr^{-1}$), and an increase in shrubland and grassland. Showing a change for more dense vegetation types. It appears to be a shift in the area of cropland from the "Centro" to the "Sur." Also, there is a high increase in forest (397 $km^2 yr^{-1}$) in the "Sur," replacing the savanna lost.

Further, we want to address whether the trend in land cover change for 2001–2023 is associated with trends in drought indices of water demand and supply and/or soil moisture for macrozone and land cover macroclasses. From the three methods tested, Ridge, Lasso, and Random Forest, neither gives significant results regarding whether the trend in a drought index for any time scale explains the trend in land cover change. Nevertheless, in "Norte Chico" and "Centro," there is a decrease in croplands and savanna and an increase in barren land, which is associated with the variation in drought indices. Mainly for a decrease in water supply (SPI and SSI) and an increase in water demand (EDDI). However, due to the high variability from north to south in Chile, the climatic conditions (arid, semi-arid, and humid), and the land cover type, we believe that only in those zones could the LULCC be driven to some degree by drought.

##	Trend of drought indices for water demand and supply, soil moisture, and vegetation productivity

```{r fig-zcNDVI_var}
#| echo: false
#| fig-pos: "!ht"
#| fig-cap: "(a) Map of the linear trend of the index zcNDVI-6 for 2001–2023. Greener colors indicate a positive trend; reder colors correspond to a negative trend and a decrease in vegetation productivity. Grey colors indicate either no vegetation or a change in land cover type for 2001–2022. (b) Temporal variation of zcNDVI-6 aggregated at macrozone level within continental Chile. Each horizontal panel corresponds to a macrozone from 'Norte Grande' to 'Austral'."
#| out-width: "\\textwidth"
knitr::include_graphics("../output/figs/temporal_variation_zcNDVI6_macrozonas_con_mapa.png")
```

In "Norte Grande," vegetation productivity, as per the z-index, exhibits a yearly increase of 0.02 in the five land cover macroclasses, with respect to grassland and shrubland categories. There is a negative trend in "Norte Chico" with -0.04 and "Centro" with -0.02 per decade. In the "Norte Chico," savanna (-0.05) has the lowest trend, and the rest of the types are around -0.04. In "Centro," shrubland reaches -0.06, grassland -0.05, and croplands and savanna -0.01 per decade. This could be associated either with a reduction in vegetation surface, a decrease in biomass, or browning [@Miranda2023]. Vegetation reached its lowest values since the year 2019, reaching an extreme condition in early 2020 and 2022 in the "Norte Chico" and Centro" (Mega Drought). The "Sur" and "Austral" show a positive trend of around 0.016 per decade (@fig-zcNDVI_var). Despite the croplands suffering from drought just as badly as the native vegetation in "Norte Chico," the Chilean matorral [@Fuentes2021] appears to be the region most affected by a negative trend in vegetation productivity.

```{r fig-trendDIMacro}
#| echo: false
#| fig-pos: "!ht"
#| fig-cap: "Trend per decade for the drought indices SPI, EDDI, SPEI, and SSI aggregated by macrozone."
#| out-width: "\\textwidth"
knitr::include_graphics('../output/figs/trend_macrozone_drought_indices.png')
```

\blandscape
```{r fig-trendDI}
#| echo: false
#| layout-ncol: 2
#| fig-pos: "!ht"
#| fig-env: "figure*"
#| fig-cap: "Linear trend of the drought index (*) at time scales of 1, 3, 6, 12, 24, and 36 months for 1981-2023"
#| fig-subcap: 
#| - "SPI (Standardized Precipitation Index)"
#| - "SPEI (Standardized Precipitation Evapotranspiration Index)" 
#| - "EDDI (Evaporative Demand Drought Index)"
#| - "SSMI (Standardized Soil Moisture Index)"
#| out-width: "\\textwidth"

knitr::include_graphics("../output/figs/trend_raster_SPI_1981-2023.png")
knitr::include_graphics("../output/figs/trend_raster_SPEI_1981-2023.png")
knitr::include_graphics("../output/figs/trend_raster_EDDI_1981-2023.png")
knitr::include_graphics("../output/figs/trend_raster_zcSM_1981-2023.png")

```

\elandscape

```{=latex}
\begin{table}[!ht]
\caption{Summarry per land cover macroclass and macrozone regarding the correlation between zcNDVI with the drought indices EDDI, SPI, SPEI, and SSI for time scales of 1, 3, 6, 12, 24, and 36. The numbers in each cell indicate the time scale that reached the maximum correlation for the land cover and macrozone, and the color indicates the strength of the r-squared obtained with the index and the time scale.}
\label{tab-corlandcover}
\includegraphics[]{../output/figs/tabla_r_cor_macro_indice.png}
\end{table}
```

Analyzing the water supply, the macrozones that have the lowest trend are "Norte Chico" and "Centro," where the SPI, SPEI, and SSI show that it decreases at longer time scales due to the prolonged reduction in precipitation. At 36 months, it reaches trends between -0.03 and -0.04 (z-score) per decade for SPI, SPEI, and SSI (@fig-trendDI). For "Sur," the behavior is similar, decreasing at longer scales and having between -0.016 and -0.025 per decade for SPI, SPEI, and SSI. On the other hand, all macrozones show an increase in the trend in all the drought indices, with "Norte Grande" having the highest at 36 months (0.042 per decade). Because of this, the SPEI (which uses AED) reached its lowest value in "Norte Grande," with -0.03 at 36 months. Despite the other macrozones, "Austral" showed an increase in all indices, being the highest for EDDI at 36 months (0.025) and the lowest for SSI, which shows only a minor increase in the trend (@fig-trendDI and @fig-trendDIMacro).

##	Impact for water supply and demand, and soil moisture in vegetation productivity

According to what is shown in @fig-corTimeScale, @fig-corPerson, and Table \ref{tab-corlandcover}, forest seems to be the most resistant type to drought. Showing that only "Centro" is slightly (rsq = 0.25) impacted by a 12-month soil moisture deficit (SSI-12). In the "Norte Chico" and to a lesser extent in the "Norte Grande," it is evident that a SSI-12 with a rsq = 0.45 and a decrease in water supply (SPI-36 and SPEI-24 with rsq = 0.28 and 0.34, respectively) have an impact on grasslands. However, this type was unaffected by soil moisture, water supply, or demand in macrozones further south. The types that show to be most affected by variation in climate conditions are shrublands, savannas, and croplands. For savannas in "Norte Chico," the SSI-12 and SPI-24 reached an rsq of 0.74 and 0.58, respectively. This value decreases to the south, but the SSI-12 is still the variable explaining more of the variation in vegetation productivity (rsq = 0.45 in "Centro" and 0.2 in "Sur"). In the case of croplands, the SPEI-12, SPI-36, and SSI-12 explain between 45% and 66% of "Norte Chico." The type of land most impacted by climatic variation was shrubland, where soil moisture explained 59% and precipitation, 37%, in "Norte Chico" and "Centro," with SSI-12 being the most relevant variable, then SPI-36 in "Norte Chico" and SPI-24 in "Sur."

```{r fig-corTimeScale}
#| echo: false
#| fig-pos: "!ht"
#| out-width: "\\textwidth"
#| fig-cap: "Time scales per drought index that reach the maximum coefficient of determination"
knitr::include_graphics("../output/figs/mapa_cor_selec_indices_zcNDVI6.png")
```

```{r fig-corPerson}
#| echo: false
#| fig-pos: "!ht"
#| out-width: "\\textwidth"
#| fig-cap: "Pearson correlation value for the time scales and drought index that reach the maximum coefficient of determination"
knitr::include_graphics("../output/figs/mapa_cor_r_indices_zcNDVI6.png")
```

#	Discussion

## Drought trend and attribution to LULCC

@Vicente-Serrano2021, in a study at the global scale of drought trends, indicates that there have not been significant trends in meteorological drought since 1950. Also, state that the increase in hidrological trend in some parts of the globe (northeast Brazil and the Mediterranean region) is related to changes in land cover and specifically to the rapidly increasing irrigated area, which consequently increases water extraction. @Kogan2020 analyzed the agricultural drought impact globally and in the main grain producer countries, finding that _"since 1980, the Earth warming has not changed the drought area or intensity"_.

In our study, we considered the variation in vegetation productivity in Chile for areas without changes in land cover macroclasses (see @sec-persistence_mask), to avoid misleading conclusions that could be related to the increase in water demand due to LULCC. Our results show a contrasting perspective. There has been a significant trend in the decline of vegetation productivity (zcNDVI) since 2000 for "Norte Chico" and "Centro," which has been extreme between 2020 and 2022, seemsly due to an intense hydrological drought due to the persistance of the Mega Drought [@Garreaud2017]. Despite using the persistance mask for vegetation's trend analysis, cropland, which is the most water-demand type, showed a decrease trend in "Norte Chico" and "Centro." Also, there was an increase in barren land for both types. These changes are associated with a decrease in water demand from vegetation. Nonetheless, we used the persistant land cover to ensure that the pixel has the same class; in the case of croplands, it could happen that some areas had changed crops for others with higher water consumption. But this effect should be minor compared to the results from ladcover macroclasses.

On the other hand, for "Norte Chico" and "Centro," our results show a decrease in trends of water supply (SPI and SSI), which are higher at larger time scales and consequently impact the hydrological system. We claim that what occurred in central Chile defies findings made at the global level [@Vicente-Serrano2021;@Kogan2020], demonstrating that a constant decrease in water supply rather than an increase in water demand (i.e., irrigated crops) is the main cause of the hydrological drought. Finally, central Chile has shown a diminishment in vegetation productivity for all macroclasses, mainly attributed to variation in water supply, i.e., precipitation, which could be strengthened by an increase in water demand by, for example, an increase in the surface area of irrigated crops.

## Land cover types and their impact by drought

We found that shrubland, savannas (Chilean matorral), and croplands are the most sensitive to climate conditions. Being most affected by the 12-month soil moisture deficit. In a study in the Yangtze River Basin in China, Jiang2020 analyzed the impact of drought on vegetation using the SPEI and the Enahanced Vegetation Index (EVI). They found that cropland was more sensitive to drought than cropland, showing that cropland responds strongly to short- and medium-term drought (< SPEI-6). In our case, the SPEI-12 was the one that most impacted the croplands in "Norte Chico" and "Centro." In general, most studies show that croplands are most sensitive to short-term drought (< SPI-6) [@Zambrano2016; @Potopova2015; @Dai2020;@Rhee2010]. Short-term precipitation deficits impact soil water, and thus less water is available for plant growth. However, we found that in "Norte Chico," an SPI-36 and SPEI-12 had a higher impact, which are associated with hydrological drought (long-term), and in "Centro," an SPI-12 and SPEI-12. Thus, we attribute this impact to the hydrological drought that has decreased groundwater storage [@Taucare2024], which in turn is impacted by long-term deficits, and consequently, the vegetation is more dependent on groundwater. In "Sur" and "Austral," the correlations between drought indices and vegetation productivity decrease, as do the time scales that reach the maximum r-squared (\ref{tab-corlandcover}). What can be explained is that, south of "Centro," predominate forest and grassland, the most resistant types. Also, drought episodes have been less frequent and intense. The drought episodes have had a lower impact on water availability for vegetation.

Extreme drought conditions are an important driver of tree mortality, as shown by @Senf2020 in Europe. However, we found that forest is the type of land cover macroclass less affected by variation in drought indices, being the most resistant land cover class to drought. Supporting this is @Fathi-Taperasht2022, who asserts that Indian forests are the most drought-resistant and recover rapidly. Similarly, the work of @Wu2024, who analyzed vegetation loss and recovery in response to meteorological drought in the humid subtropical Pearl River basin in China, indicates that forests showed higher drought resistance. Using Vegetation Optical Depth (VOD), kNDVI, and EVI, @Xiao2023, tests the resistance of ecosystems and finds that ecosystems with more forests are better able to handle severe droughts than croplands. They attribute the difference to a deeper rooting depth of trees, a higher water storage capacity, and different water use strategies between forest and cropland [@Xiao2023].

In contrast to what we obtained, @Venegas2022, who studied Cryptocarya alba and Beilschmiedia miersii (both from the Lauraceae family) that live in sclerophyllous forests in Chile, found that the trees' overall growth had slowed down. This could mean that the natural dynamics of their forests have changed. They attributed it to the cumulative effects of the unprecedented drought (i.e., hydrological drought). Thus, we attribute that forest to being the most resistant to drought, due to the fact that most of the species comprising it are highly resilient to water scarcity compared to the other land cover classes. Nonetheless, if we want to go deep in our analysis, we should use earth observation data that is able to capture a higher level of detail. For example, when we used MOD13A3 with a 1km spatial resolution to measure vegetation condition, it took the average condition of 1 square kilometer. Then, to study how a type of forest (e.g., sclerophyllous forest) changes in response to drought on a local level using remote sensing, we should use operational products with higher spatial resolutions, such as those from Landsat or Sentinel.

## Soil moisture, vegetation productivity, and agricultural drought.

The main external factors that affect biomass production by vegetation are ET and SM, and the rate of ET in turn depends on the availability of water storage in the root zone. Thus, soil moisture plays a key role in land carbon uptake and, consequently, in the production of biomass [@Humphrey2021]. Moreover, @Zhang2022 indicates there is a bidirectional causality between soil moisture and vegetation productivity. Lastly, some studies have redefined agricultural drought as soil moisture drought from an hydrological perspective [@Loon2016;@Samaniego2018]. Even though soil moisture is the external factor most determinant of vegetation biomass, there are multiple internal factors, such as species, physiological characteristics, and plant hydraulics, that would affect vegetation productivity. Because of that, we believe that agricultural drought, referring to the drought that impacts vegetation productivity, is the most proper term, as originally defined by @Wilhite1985.

The study results showed that the soil moisture-based drought index (SSI) was better at explaining vegetation productivity across land cover macroclasses than meteorological drought indices like SPI, SPEI, and EDDI. In the early growing season and especially in irrigated rather than rainfed croplands, soil moisture has better skills than SPI and SPEI for estimating gross primary production (GPP), according to @Chatterjee2022's evaluation of the SPI and SPEI and their correlation with GPP in the CONUS. Also, @Zhou2021 indicates that the monthly scaled Standardized Water Deficit Index (SWDI) can accurately show the effects of agricultural drought in most of China. Complementary, @Nicolai2017 analyzed the time-lag between SWDI and the Vegetation Condition Index (VCI), and they state that there was no or little time-lag with croplands but a significant time-lag in the case of forests.

In our case, there is strong spatial variability throughout Chile and between classes, mainly attributable to climate heterogeneity, hydrological status, or vegetation resistance to water scarcity. The semi-arid "Norte Chico" and the Mediterranean "Centro" were where SSI had the best performance. In Chile, medium-term deficits of 12 months are more relevant in the response of vegetation, which decreases to the south, and in the case of croplands, they seem to react in a shorter time, with six months (SSI-6) in "Centro." This variation for croplands could be related to the fact that in "Norte Chico," the majority of crops are irrigated, but to the south there is a higher proportion of rainfed agriculture, which is most dependent on the short-term availability of water. Rather, in the "Norte Chico," the orchards are more dependent on the storage of water in dams of groundwater reservoirs, which are affected by long-term drought (e.g., SPI-36).

## Early drought forecasting for vegetation productivity

We analyzed the correlation between meteorological and soil moisture drought indices with zcNDVI. From our findings, we could further use the drought indices and time scales that have the highest r-squared to develop a combined model to forecast vegetation productivity in Chile. Despite the fact that SSI was the best-performing index, the rest of the indices should help to enhance predictability. @Zambrano2018 proposed a prediction model for cropland surface in Chile at administrative units with a 1- to 6-month lead time using zcNDVI from MODIS and climate oscilation indices. The results given were r-squared, ranging from 0.95 at a 1-month lead time to 0.37 at a 6-month lead time. Thus, incorporating the results of this study with those made by @Zambrano2018, we could develop a combined forecasting model at the pixel level for the macroclasses in Chile based on drought indices of water demand and supply and soil moisture.

## Future outlook (to complete)

# Conclusion

There is a trend toward decreasing water supply in most parts of Chile, less in the "Austral," which is stronger in the "Centro" and "Norte Chico." The whole country showed an increase in water demand. Vegetation productivity only showed a decrease in the "Norte Chico" and "Centro," being highest for shrubland and croplands. Forest is the land cover most resistant to drought, and shrubland and cropland are the most sensitive.

A soil moisture deficit of 12 months (SSI-12) is highly correlated with vegetation productivity for the land cover classes of shrubland, savannas, croplands, and forest in "Norte Chico" and "Centro." For the southern part of the country with humid conditions, the correlation with SSI decreases. Soil moisture overcomes the capacity to explain vegetation productivity over the supply and demand drought indices in the entire territory.

The variation in vegetation productivity appears to be associated with climate variation rather than anthropogenic factors (e.g., an increase in water demand by irrigated crops). Even though switching to more demanding crops on the land could increase the impact of drought on vegetation, this would need to be more thoroughly investigated, for instance at the watershed level.

The results of this study could help in the development of a robust forecasting system for land cover classes in Chile., helping to improve preparedness for climate change impacts on vegetation.

# Supplementary material {-}

# References {-}