### 2024-03-19 article/paper_drought_Chile2023.qmd
update article

---
title: The effects of multi-dimensional drought on land cover change and vegetation productivity in continental Chile
#subtitle: A Short Subtitle
author:
  - name: Francisco Zambrano
    email: francisco.zambrano@umayor.cl
    affiliations: 
        - id: umayor
          name: Universidad Mayor
          department: Hémera Centro de Observación de la Tierra, Facultad de Ciencias, Escuela de Ingeniería en Medio Ambiente y Sustentabilidad
          #address: Street Address
          city: Santiago, Chile
          #state: State
          postal-code: 7500994
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Anton Vrieling
    email: a.vrieling@utwente.nl
    affiliations:
        - id: itc
          name: University of Twente
          department: Faculty of Geo-Information Science and Earth Observation (ITC)
          #address: Street Address
          city:  Enschede, The Netherlands
  - name: Francisco Meza
  - name: Iongel Duran-Llacer
  - name: Francisco Fernández
    email: francisco.fernandez@uss.cl
    affiliations:
        - id: uss
          name: Universidad San Sebastian
          department: 
          #address: Street Address
          city:  
  - name: Alejandro Venegas-González
  - name: Nicolas Raab
  - name: Dylan Craven        
  #   note: |
  #     Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
  # - name: Cat Memes
  #   email: cat@example.com
  #   affiliations:
  #       - ref: another-u
  #   note: Yet another author footnote.
  # - name: Derek Zoolander
  #   email: derek@example.com
  #   affilations:
  #       - ref: some-tech
abstract: |
 The north-central region of Chile has been the focus of research studies due to the persistent decrease in water supply, which is impacting the hydrological system and vegetation development. This persistent period of water scarcity has been defined as a megadrought. The aim of our study is to evaluate the land cover change over continental Chile and to examine how this is connected to drought indices of water supply, atmospheric evaporative demand (AED), soil moisture, and their effects on vegetation productivity. The drought indices were derived using monthly ERA5-Land reanalysis data spanning from 1981 to 2023. The Moderate-Resolution Imaging Spectroradiometer (MODIS) datasets were utilized to obtain information on annual land cover and monthly vegetation productivity. We analyzed short- (1, 3, 6 months) to long-term (12, 24, 36 months) time scales of drought. Our results showed that land cover change was highest in the south-central part of the country, reaching changes as high as 36% in the surface type. The water demand has increased for the whole country, with a major increase in the north. The AED and soil moisture evidence a decreasing trend, which decreases at longer time scales and from north to south. The extreme south part of the country shows an increase in supply. Vegetation productivity has a negative trend in the north-central region for all land cover types. On the other hand, forests seem to be the most resistant type to drought. The types that show to be most affected by variation in climate conditions are shrublands, savannas, and croplands. The drought indices that have the capability of explaining to a major degree the variance in vegetation productivity are the ones that consider soil moisture for twelve months and the combined effect of precipitation and AED for 24 and 12 months. The results indicate that the north-central region is the most sensitive to water supply deficits lasting longer than a year. 
keywords: 
  - drought
  - land cover change
  - satellite
date: last-modified
bibliography: references.bib
keep-tex: false
include-before-body:
  - text: |
      \captionsetup{justification=raggedright,singlelinecheck=false}
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      layout: onecolumn
      model: 3p
      cite-style: authoryear
      natbiboptions: numbers
    include-in-header: 
      text: \usepackage{lineno}\linenumbers
            \usepackage{multirow}
            \usepackage{lscape}
            \newcommand{\blandscape}{\begin{landscape}}
            \newcommand{\elandscape}{\end{landscape}}
  docx:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      layout: onecolumn
      model: 3p
      cite-style: authoryear
      natbiboptions: numbers
    include-in-header: 
      text: \usepackage{lineno}\linenumbers
            \usepackage{multirow}
            \usepackage{lscape}
            \newcommand{\blandscape}{\begin{landscape}}
            \newcommand{\elandscape}{\end{landscape}}
---

# Introduction

Drought is often classified as meteorological when there is a decrease in precipitation below the mean average of several years (more than 30 years), hydrological when these anomalies last for long periods (months to years) and affect water systems, and agricultural when the deficit impacts plant health anomalies and leads to decreased productivity [@Wilhite1985]. However, it is important to note that drought is also influenced by human activities, which were not considered in the definitions. Thus, @Loon2016 and @AghaKouchak2021 have given an updated definition of drought for the Anthropocene, suggesting that it should be considered the feedback of humans’ decisions and activities that drives the anthropogenic drought. Simultaneously, drought leads to heightened tree mortality and induces alterations in land cover and land use, ultimately affecting ecosystems [@Crausbay2017]. Even though many ecological studies have misinterpreted how to characterize drought, for example, sometimes considering "dry" conditions as "drought" [@Slette2019]. Then, @Crausbay2017 proposed the ecological drought definition as "an episodic deficit in water availability that drives ecosystems beyond thresholds of vulnerability, impacts ecosystem services, and triggers feedback in natural and/or human systems." In light of current global warming, it is crucial to study the interaction between drought and ecosystems in order to understand their feedback and impact on water security. [@Bakker2012]

Human-induced greenhouse gas emissions have increased the frequency and/or intensity of drought as a result of global warming, according to the sixth assessment report (AR6) of the Intergovernmental Panel on Climate Change (IPCC) [@IPCC2023]. The evidence supporting this claim has been strengthened since AR5 [@IPCC2013]. Recent studies, however, have produced contrasting findings, suggesting that drought has not exhibited a significant trend over the past forty years. [@Vicente-Serrano2022; @Kogan2020]. @Vicente-Serrano2022 analyzed the meteorological drought trend on a global scale, finding that only in a few regions has there been an increase in the severity of drought. Moreover, they attribute the increase in droughts over the past forty years solely to an increase in atmospheric evaporative demand (AED), which in turn enhances vegetation water demand, with important implications for agricultural and ecological droughts. Also, they state that "the increase in hydrological droughts has been primarily observed in regions with high water demand and land cover change". Similarly, @Kogan2020 analyzed the drought trend using vegetation health methods, finding that for the globe, hemispheres, and main grain-producing countries, drought has not expanded or intensified for the last 38 years. Further, the @IPCC2021 suggests that there is a high degree of confidence that rising temperatures will increase the extent, frequency, and severity of droughts. Also, AR6 [@IPCC2023] predicts that many regions of the world will experience more severe agricultural and ecological droughts even if global warming stabilizes at 1.5°–2°C. To better evaluate the impact of drought trends on ecosystems, assessments are needed that relate meteorological and soil moisture variables to their effects on vegetation.

From 1960 to 2019, land use change has impacted around one-third of the Earth's surface, which is four times more than previously thought [@Winkler2021]. Multiple studies aim to analyze and forecast changes in land cover globally [@Winkler2021; @Song2018] and regionally [@Chamling2020; @Homer2020; @Yang2021]. Some others seek to analyze the impact of land cover change on climate conditions such as temperature and precipitation [@Luyssaert2014;@Pitman2012]. There is less research on the interaction between drought and land cover change [@Chen2022;@Akinyemi2021;@Peng2017]. @Peng2017 conducted a worldwide investigation utilizing net primary production to examine the spatial and temporal variations in vegetation productivity at global level. The study aimed to assess the influence of drought by comparing the twelve-month Standardized Precipitation Evapotranspiration Index (SPEI) and land cover change. According to their findings, drought is responsible for 37% of the decline in vegetation productivity, while water availability accounts for 55% of the variation. @Chen2022 studied the trend of vegetation greenness and productivity and its relation to meteorological drought (SPEI of twelve months in December) and soil moisture at the global level. The results showed lower correlations (<0.2) for both variables. @Akinyemi2021 evaluates drought trends and land cover change using vegetation indices in Botwsana in a semi-arid climate. These studies mostly looked at how changes in land cover and vegetation productivity are related to a single drought index (SPEI) over a single time period of 12 months. SPEI takes into account the combined effect of precipitation and AED as a water balance, but it does not allow us to know the contribution of each variable on its own. Some things worth investigating in terms of land cover change and vegetation productivity are: i) How do they respond to short- to long-term meteorological and soil moisture droughts? ii) How is the drought impacting land cover changes? And iii) How do they behave in humid and arid climatic zones regarding drought? Likewise, there is a lack of understanding of how the alteration in water supply and demand is affecting land cover transformations.

For monitoring drought, the World Meteorological Organization recommends the SPI (Standardized Precipitation Index) [@WMO2012]. The SPI is a multi-scalar drought index that only uses precipitation to assess short- to long-term droughts. The primary cause of drought is precipitation anomalies, and temperature usually makes it worse [@Luo2017]. Nowadays, there is an increase in attention toward using AED separately to monitor droughts [@Vicente-Serrano2020]. One reason is due to its attribution to increasing flash droughts in water-limited regions [@Noguera2022]. @Vicente-Serrano2010 proposed the Standardized Precipitation Evapotranspiration Index (SPEI), which incorporated the temperature effect by subtracting AED from precipitation. SPEI allows for analysis of the combined effect of precipitation and AED. Since its formulation, it has been used worldwide for the study and monitoring of drought [@Gebrechorkos2023;@Liu2021]. @Hobbins2016 and @McEvoy2016 developed the Evaporative Demand Drought Index (EDDI) to monitor droughts solely using the AED, and it has proven effective in monitoring flash droughts [@Li2024;@Ford2023]. For soil moisture, several drought indices exist, such as the Soil Moisture Deficit Index (SDMI) [@Narasimhan2005] and the Soil Moisture Agricultural Drought Index (SMADI) [@Souza2021]. @Hao2013 and @AghaKouchak2014 proposed the Standardized Soil Moisture Index (SSI), which has a similar formulation as the SPI, SPEI, and EDDI. Thus, there are plenty of drought indices that allow for a comprehensive assessment of drought from short- to long-term scales and consider variables from the earth's water balance on their own (e.g., precipitation, AED, soil moisture). Using this information, we can advance our understanding of the impact of drought on ecosystems.

Chile's diverse climatic and ecosystem types [@Beck2023;@Luebert2022] make it an ideal natural laboratory for studying climate and ecosystems. Additionally, the country has experienced severe drought conditions that have had significant effects on vegetation and water storage. North-central Chile has faced a persistent precipitation deficit since 2010, defined as a mega drought. [@Garreaud2017], which has impacted the Chilean ecosystem. This megadrought was defined by the Standardized Precipitation Index (SPI) of twelve months in December having values below one standard deviation. Some studies have addressed how this drought affects single ecosystems in terms of forest development [@Miranda2020;@Venegas2018], forest fire occurrence [@UrrutiaJalabert2018], and crop productivity [@Zambrano2023;@Zambrano2018;@Zambrano2016]. We found one study regarding land cover and drought in Chile. The study by @Fuentes2021  evaluates water scarcity and land cover change in Chile between 29° and 39° of south latitude. @Fuentes2021 used the SPEI of one month for evaluating drought, which led to misleading results. For example, they did not find a temporal trend in the SPEI but found a decreasing trend in water availability and an increase trend on AED, which in turn should have been capable of being captured with longer time scales of the SPEI. The term "megadrought" in Chile is used to describe a prolonged water shortage that lasts for several years, resulting in a permanent deficit that impacts the hydrological system [@Boisier2018]. Therefore, it is crucial to evaluate temporal scales that consider the cumulative impact over a period of several years. There is limited understanding of the correlation between drought and the environment in Chile. Therefore, it is crucial to have a deeper understanding of how meteorological and soil moisture droughts impact the dynamics of the environment in order to make informed decisions on how to adapt.

<!-- A detailed spatiotemporal assessment of the interaction of drought for short- to long-term and land cover change requires information on vegetation as well as weather variables such as precipitation, temperature, and soil moisture. Weather networks in Chile present some disadvantages, such as spatio-temporal gaps, a short history, and irregular quality, which make them difficult to represent the whole extent of the country spatially. In order to do this, we use reanalysis data from ERA5-Land [@MunozSabater2021] to create drought indices that consider AED, precipitation, and soil moisture over a range of time periods, from the short to the long term. Also, we use vegetation spectral information and annual land cover change from the Moderate-Resolution Imaging Spectroradiometer (MODIS). We expect to gain insight regarding the temporal evolution of water demand, water supply, and soil moisture, as well as the interaction with land cover change and vegetation productivity.  -->

Here, we analyze the multi-dimensional impacts of drought across ecosystems in continental Chile. More specifically, we aim to assess: i) short- to long-term temporal trends in multi-scalar drought indices; ii) temporal changes in land-use cover and the direction and magnitude of their relationships with trends in drought indices; and iii) the trend in vegetation productivity and its relationship with drought indices across Chilean ecosystems.

# Study area

Continental Chile has diverse climate conditions with strong gradients from north to south and east to west [@Aceituno2021] (@fig-studyArea a), which determines its great ecosystem diversity [@Luebert2022] (@fig-studyArea c). The Andes Mountains are a main factor in climate latitudinal variation [@Garreaud2009]. "Norte Grande" and "Norte Chico" predominate in an arid desert climate with hot (Bwh) and cold (Bwk) temperatures. At the south of "Norte Chico," the climate changes to an arid steppe with cold temperatures (Bsk). In these two northern regions, the land is mostly bare, with a small surface of vegetation types such as shrubland and grassland. In the zones "Centro" and the north half of "Sur," the main climate is Mediterranean, with warm to hot summers (Csa and Csb). Land cover in "Centro" comprises a significant amount of shrubland and savanna (50%), grassland (16%), forest (8%), and croplands (5%). An oceanic climate (Cfb) predominates in the south of "Sur" and the north of "Austral." Those zones are high in forest and grassland. The southern part of the country has a tundra climate, and in "Austral," it is a cold semi-arid area with an extended surface of grassland, forest, and, to a lesser extent, savanna.

```{r fig-studyArea}
#| echo: false
#| fig-pos: "!ht"
#| fig-env: 'figure*'
#| fig-cap: '(a) Chile with the Koppen-Geiger climate classes and the five macrozones "Norte Grande", "Norte Chico",  "Centro", "Sur", and "Austral". (b) Topography reference map. (c) land cover classes for 2022. (d) Persistent land cover classes (> 80%) for 2001-2022'
#| out-width: "\\textwidth"
knitr::include_graphics("../output/figs/map_study_con_landcover.png")
```

# Materials and Methods

##	Data

### Gridded meteorological and vegetation data

To analyze land cover change, we use the classification scheme by the IGBP (International Geosphere-Biosphere Programme) from the product MCD12Q1 collection 6.1 from MODIS. To derive a proxy for vegetation productivity, we used the Normalized Difference Vegetation Index (NDVI) from the product MOD13A3 collection 6.1 from MODIS [@Didan2015]. MOD13A3 provides vegetation indices at 1km of spatial resolution and monthly frequency. The NASA EOSDIS Land Processes Distributed Active Archive Center (LP DAAC), USGS Earth Resources Observation and Science (EROS) Center, Sioux Falls, South Dakota, provided the MOD13A3 and MCD12Q1 from the online Data Pool, accessible at https://lpdaa.usgs.gov/tools/data-pool/.

```{=latex}
\begin{table}[!ht]
\caption{Description of the satellite and reanalysis data used}
\label{tab-desEOD}
\small
\centering
\begin{tabular}{p{0.13\textwidth}cp{0.3\textwidth}p{0.095\textwidth}ccc}
\hline
\multirow{1}{*}{\centering Product} & Sub-product & Variable & Spatial Resolution  & Period & Units & Short Name \\ 
\hline
\multirow{4}{*}{ERA5L} & ~ & Precipitation & \multirow{4}{*}{~0.1°} & \multirow{4}{*}{1981-2023} & mm & P \\ 
         &  & Maximum temperature & ~ & & $°C$ & $T_{max}$ \\ 
         &  & Minimum temperature & ~ & & $°C$ & $T_{min}$ \\ 
         &  & Volumetric Soil Water Content at 1m & ~ & & $m3/m3$ & SM \\ 
ERA5L* & & Atmospheric Evaporative Demand & 0.1° & 1981-2023 & mm & AED \\
        \multirow{2}{*}{MODIS} & MOD13A3.061 & Normalized Difference Vegetation Index & \multirow{2}{*}{~1 km} & 2000-2023 & ~ & NDVI \\ 
         & MCD12Q1.061 & land cover IGBP scheme & & 2001-2022 & ~ & land cover \\ 
\hline
\end{tabular}
{\raggedright *Calculated from maximum and minimum temperatures derived from ERA5L with Eq. \ref{eq-AED}. \par}
\end{table}
```

For soil moisture, water supply, and water demand variables, we used ERA5L (ECMWF Reanalysis version 5 over land) [@MunozSabater2021], a reanalysis dataset that provides the evolution of atmospheric and land variables since 1950. It has a spatial resolution of 0.1° (9 km), hourly frequency, and global coverage. We selected the variables for total precipitation, maximum and minimum temperature at 2 meters, and volumetric soil water layers between 0 and 100cm of depth (layer 1 to layer 3). Table \ref{tab-desEOD} shows a summary of the data and its main characteristics.

## Trend of short- to long-term drought

### Atmospheric Evaporative Demand (AED)

In order to compute the drought indices that use water demand, it is necessary to first calculate the AED. To do this, we employed the Hargreaves method [@Hargreaves1994;@Hargreaves1985] by applying the following equation:

$$AED = 0.0023\cdot Ra\cdot (T+17.8)\cdot (T_{max}-T_{min})^{0.5}$$ {#eq-AED}

where $Ra$ $(MJ\,m^2\, day^{-1})$ is extraterrestrial radiation; $T$, $T_{max}$, and $T_{min}$ are mean, maximum, and minimum temperature $(°C)$ at 2m. For calculating $Ra$ we used the coordinate of the latitud of the centroid of each pixel. We chose the method of Hargreaves to estimate AED because of its simplicity, which only requires temperatures and extrarrestrial radiation. Also, it has been recommended over other methods (e.g., Penman-Monteith) when the access to climatic variables is limited [@Vicente-Serrano2014].

### Non-parametric calculation of drought indices

To derive the drought indices of water supply and demand, soil moisture, and vegetation (i.e., the proxy of productivity), we used the ERA5L dataset and the MODIS product, with a monthly frequency for 1981–2023 and 2000–2023, respectively.

The dought indices correspond to a historical anomaly with regard to a variable (e.g., meteorological, vegetation, or soil moisture). To account for the anomaly, the common practice is to derive it following a statistical parametric methodology in which it is assumed that the statistical distribution of the data is known (@Heim2002). A wrong decision is usually the highest source of uncertainty (@Laimighofer2022). In the case of Chile, due to its high degree of climatic variability, it is complex to choose a proper distribution without previous research. Here, we follow a non-parametric methodology for the calculation of the drought indices, in a similar manner as the framework proposed by @Farahmand2015; @Hobbins2016;@McEvoy2016.

For the purpose of monitoring water supply drought, we used the well-known Standardized Precipitation Index (SPI), which the World Meteorological Organization (WMO) recommended. The SPI solely relies on precipitation data. Also, it has been used worldwide for the study of drought, including in Chile (@Garreaud2017; @Zambrano2017). The primary cause of drought is precipitation anomalies, and temperature usually makes it worse [@Luo2017]. Nowadays, there is an increase in attention toward using water demand separately to monitor droughts. (@Vicente-Serrano2020; @Noguera2022). Thus, to evaluate water demand, we chose the Evaporative Demand Drought Index (EDDI), developed by @Hobbins2016 and @McEvoy2016, which is based on the AED. EDDI is currently used for monitoring drought in the United States (https://psl.noaa.gov/eddi/). In our case, we used only temperature for AED, a difference from the original formulation of EDDI, which also considered wind besides temperature. To consider the combined effect of water supply and demand, we selected the SPEI, which corresponds to a balance between precipitation and AED. @Vicente-Serrano2010 proposed the SPEI, and it has improved the SPI by incorporating temperature for drought monitoring. For SPEI, an auxiliary variable D = P-AED is calculated. Soil moisture is the main driver of vegetation productivity, particularly in semi-arid regions (@Li2022). Hence, for soil water drought, we used the SSI (Standardized Soil Moisture Index) [@Hao2013;@AghaKouchak2014] which is a multi-scale index similar to SPI, SPEI, and EDDI. In our case, for the SSI, we used the average soil moisture from ERA5L at 1m depth. Finally, for the proxy of productivity, we used the zcNDVI proposed by @Zambrano2018 which will be derived from the NDVI retrieved from MOD13A1.

To derive the drought indices, first we must calculate the sum of the variables with regard to the time scale (s). In this case, for generalization purposes, we will use $V$, referring to variables $P$, $AED$, $D$, $NDVI$, and $SM$ (Table \ref{tab-desEOD}). We cumulated each $V$  over the time series of $n$ values (months), and for the time scales $s$:

$$A_{si} = \sum_{i=n-s-i+2}^{n-i+1} V_i\,\, \forall\, i\geq n-s+1  $$ {#eq-sumvar}

The $A_{si}$ corresponds to a moving window (convolution) that sums the variable for time scales $s$ from the last month, month by month, until the first month in which it could sum for $s$ months. Once the variable is cumulated over time for the scale $s$. Thus, the empirically derived probabilities are obtained through an inverse normal approximation [@Abramowitz1968]. Then, we used the empirical Tukey plotting position [@Wilks2011] over $A_i$ to derive the $P(a_i)$  probabilities across a period of interest:

$$P(A_i) = \frac{i-0.33}{n+0.33'}$$ {#eq-probPai}

The drought indices $SPI$, $SPEI$, $EDDI$, $SSI$, and $zcNDVI$ are obtained following the inverse normal approximation:

$$DI(A_i) = W - \frac{C_0+C_1\cdot W + c_2 \cdot W^2}{1+d_1\cdot W +d_2\cdot W^2 +d_3\cdot W^3}$$ {#eq-DI}

$DI$ is referring to the drought index calculated for the variable $V$. The values for the constats are: $C_0 = 2.515517$, $C_1 = 0.802853$, $C_2 = 0.010328$, $d_1 = 1.432788$, $d_2 = 0.189269$, and $d3 = 0.001308$. For $P(A) \leq 0.5$, W=$\sqrt{-2\cdot ln(P(A_i))}$ , and for $P(A_i) > 0.5$, replace $P(A_i)$ with $1-P(A_i)$ and reverse the sign of $DI(A_i)$.

The drought indices were calculated for time scales of 1, 3, 6, 12, 24, and 36 months at a monthly frequency for 1981–2023 in order to be used for short- to long-term evaluation of drought. In the case of the proxy of vegetation productivity (zcNDVI) it was calculated for a time scale of six months at monthly frequency for 2000–2023. For zcNDVI, we test time scales of 1, 3, 6, and 12 months; we choose to use six months because that shows a more robust representation of vegetation productivity due to the seasonality of vegetation in Chile.

### Trend of drought indices

To estimate if there are significant positive or negative trends for the drought indices, we used the non-parametric test of Mann-Kendall [@Kendall1975]. To determine the magnitude of the trend, we used Sen’s slope [@Sen1968]. Some of the advantages of applying this methodology are that the Sen’s slope is not affected by outliers as regular regression does, and it is a non-parametric method that is not influenced by the distribution of the data. We applied Mann-Kendall to see if the trend was significant and Sen's slope to estimate the magnitude of the trend. We did this to the six time scales from 1981 to 2023 (monthly frequency) and the indices SPI, EDDI, SPEI, and SSI. Thus, we have trends per index and time scale (24 in total). Then, we extracted the trend aggregated by macrozone and per land cover persitent macroclasses.

## Interaction of land cover and drought

### Land cover change

To analyze the land cover change, we use the IGBP scheme from the MCD12Q1 collection 6.1 from MODIS. @Zambrano2018 and @Fuentes2021 have previously used this product for studies of drought and land cover in Chile. The MCD12Q1 has a yearly frequency from 2001 to 2022. The IGBP defines 17 classes; from these, we regrouped into ten macroclasses, as follows: classes 1-4 to forest, 5-7 to shrublands, 8-9 to savannas, 10 as grasslands, 11 as wetlands, 12 and 14 to croplands, 13 as urban, 15 as snow and ice, 16 as barren, and 17 to water bodies. Thus, we have a land cover raster time series with the ten macroclasses for 2001 and 2023. We validate the land cover macroclasses regarding a highly detailed (30 m of spatial resolution) land cover map made for Chile by @Zhao2016 for 2013-2014. Our results showed a global accuracy of ~0.82 and a F1 score of ~0.66. Section S2 in the Supplementary Material shows the procedure for validation.

Climate, vegetation development, seasonality, and changes in vegetation type all have an impact on the time series of NDVI. In this study, we want to examine the variation in vegetation productivity across various land cover types and how water demand, water supply, and soil moisture affect it. In order to avoid changes due to a change in the land cover type that will wrongly impact NDVI, we developed a persistence mask for land cover for 2001–2022. Thereby, we reduce an important source of variation on a regional scale. Therefore, we generated a raster mask for IGBP MODIS per pixel using macroclasses that remained unchanged for at least 80% of the years (2001–2022). This enabled us to identify regions where the land cover macroclasses are persistent.
We calculated the surface occupied per land cover class into the five macrozones ("Norte Grande" to "Austral") per year for 2001–2023. After that, we calculated the trend’s change in surface per type. We used the Sen’ slope [@Sen1968] based on Mann-Kendall [@Kendall1975].

### Relationship between land cover and drought trends

We wanted to explore the relationship between the trend in land cover classes and the trend in the drought indices. For this purpose, in order to have more representative results, we conducted the analysis over sub-basins within continental Chile. We use 469 basins, which have a surface area between 0.0746 and  24,000 ($km^2$), and a median area of 1,249 ($km^2$). For each basin, we calculate the relative trend per land cover type, considering the proportion of the type relative to the total surface of the basin. Then, we extracted per basin the average trend of the drought indices SPI, SPEI, EDDI, SSI, and all their time scales 1, 3, 6, 12, 24, and 36. Also, we extracted the average trend in the proxy of vegetation productivity (zcNDVI). We wanted to analyze which drought indices and time scales have a major impact on changes in land cover type.

We have 25 predictors, which are drought indices and vegetation productivity. We analyzed the 25 predictors per type of landcover. For the analysis, we selected the method of random forest (@Ho1995).  Because it allows to find no linear relationship, it reduces overfitting and can derive the feature importance, which helps for a better understanding of the relationships. The importance of the variable is calculated by permuting out-of-bag (OOB) data per tree and calculating the mean standard error in the OOB. Then the same is done after permuting each predictor variable. Random forest uses multiple decision trees and allows for classification and regression.

We analyzed the 25 predictors per type of landcover, thus running six models. We used random forests for regression and trained 1000 forests. For more reliable results for the important variables, we resampled by creating ten folds, running a random forest per fold, and calculating the r-squared (rsq), root mean square error (RMSE), and variable importance ten times.

## Drought impacts on vegetation productivity

We analyzed the trend of vegetation productivity over the unchanged land cover macroclasses. This way, we tried to reduce the noise in the vegetation due to a change in land cover from year to year. To achieve this, we will use the persistent mask of land cover macroclasses, which are the types that remain more than 80% of the time for 2001–2022. We used this to evaluate the trend in zcNDVI per land cover class and macrozone.

We examine the drought indices of water demand, water supply, soil moisture, and their connection with vegetation productivity to investigate two main questions: i) whether short-term or long-term time scales have a greater impact on vegetation across Chile and its specific regions; and ii) the spatial variation in the strength of the correlation between the variables and time scales.  Then, we will summarize for each land cover class and macrozone. Thus, we will be able to advance in understanding how climate is affecting vegetation, considering the impact on the five macroclasses of vegetation: forest, cropland, grassland, savanna, and shrubland.

We conducted an analysis on the linear correlation between the indices SPI, SPEI, EDDI, and SSI over time periods of 1, 3, 6, 12, 24, and 36 months, and zcNDVI. The objective is to determine the impact of soil moisture and water demand and supply on vegetation productivity. We used a method similar to that used by @Meroni2017 which compared the SPI with the cumulative FAPAR (Fraction of Absorbed Photosynthetically Active Radiation). A pixel-to-pixel linear correlation analysis was performed for each index within the persistent mask of land cover macroclasses. To begin, the Pearson coefficient of correlation is computed for each of the six time scales. A significant time scale is identified as the one that attains the highest correlation (p < 0.05). Subsequently, the Pearson correlation coefficient corresponding to the time scales at which the value peaked was extracted. As a result, for each index, we generated two raster maps: 1) containing the raster with values of the time scales that reached the maximum correlation, and 2) having the value of the correlation obtained.

## Software and packages used

For the downloading, processing, and analysis of the spatio-temporal data, we used the open source software for statistical computing and graphics, `R` [@R2023]. For downloading ERA5L, we used the `{ecmwfr}` package [@Hufkens2019]. For processing raster data, we used `{terra}` [@Hijmans2023] and `{stars}` [@Pebesma2023]. For managing vectorial data, we used `{sf}` [@Pebesma2018]. For the calculation of AED, we used `{SPEI}` [@Bergueria2023]. For mapping, we use {tmap} [@Tennekes2018]. For data analysis, the suite {tidyverse} [@Wickham2019] was used.

#	Results

## Trend of short- to long-term drought

@fig-trendDI shows the spatial variation of the trend for the drought indices from short- to long-term scales. SPI and SPEI have a decreasing trend from "Norte Chico" to "Sur." However, there is an increasing trend in "Austral." The degree of the trend is stronger at higher time scales. The SSI indicates that in "Norte Grande," there are surfaces that have increased in the soutwest part and in the northeast have decreased, and is shown for all time scales. Similar to SPI and SPEI, SSI decreases at higher time scales. EDDI showed a positive trend for the whole of continental Chile, with a higher trend toward the north and a descending gradient toward the south. The degree of trend increases at higher time scales.

\blandscape
```{r fig-trendDI}
#| echo: false
#| layout-ncol: 2
#| fig-pos: "!ht"
#| fig-env: "figure*"
#| fig-cap: "Linear trend of the drought index (*) at time scales of 1, 3, 6, 12, 24, and 36 months for 1981-2023"
#| fig-subcap: 
#| - "SPI (Standardized Precipitation Index)"
#| - "SPEI (Standardized Precipitation Evapotranspiration Index)" 
#| - "EDDI (Evaporative Demand Drought Index)"
#| - "SSMI (Standardized Soil Moisture Index)"
#| out-width: "\\textwidth"

knitr::include_graphics("../output/figs/trend_raster_SPI_1981-2023.png")
knitr::include_graphics("../output/figs/trend_raster_SPEI_1981-2023.png")
knitr::include_graphics("../output/figs/trend_raster_EDDI_1981-2023.png")
knitr::include_graphics("../output/figs/trend_raster_zcSM_1981-2023.png")

```

\elandscape

```{r fig-trendDIMacro}
#| echo: false
#| fig-pos: "!ht"
#| fig-cap: "Trend per decade for the drought indices SPI, EDDI, SPEI, and SSI aggregated by macrozone."
#| out-width: "\\textwidth"
knitr::include_graphics('../output/figs/trend_macrozone_drought_indices.png')
```

In @fig-trendDIMacro, the averaged aggregation per macrozone, drought index, and time scale are shown. The macrozones that have the lowest trend are "Norte Chico" and "Centro," where the SPI, SPEI, and SSI show that it decreases at longer time scales. Potentially explained due to the prolonged reduction in precipitation that has affected the hydrological system in Chile. At 36 months, it reaches trends between -0.03 and -0.04 (z-score) per decade for SPI, SPEI, and SSI. For "Sur," the behavior is similar, decreasing at longer scales and having between -0.016 and -0.025 per decade for SPI, SPEI, and SSI. "Norte Grande" has the highest trend at 36 months for EDDI (0.042 per decade), and "Centro" has the lowest for SPI and SPEI. In "Norte Grande" and "Norte Chico," which are in a semi-arid climate, it is evident that the EDDI has an effect on the difference between the SPI and SPEI index, which is not seen in the other macrozones. Contrary to the other macrozones, “Austral” showed an increase in all indices, being the highest for EDDI at 36 months (0.025) and the lowest for SSI, which shows only a minor increase in the trend.

### Interaction of land cover and drought

## Land cover change

```{=latex}
\begin{table}[!ht]
\caption{Surface of the land cover class that persist during 2001-2022}
\label{tab-landcoverSurf}
\includegraphics[width = .5\textwidth]{../output/figs/table_surface_landcover_macrozone.png}
\end{table}
```

```{r fig-LCprop}
#| echo: false
#| fig-pos: "!ht"
#| fig-cap: "Proportion of land cover class from the persistent land cover for 2001-2022 (>80%) per macrozone"
#| out-width: "\\textwidth"
knitr::include_graphics("../output/figs/LC_pers80_per_macrozone.png")
```

For vegetation, we obtained and use hereafter five macroclasses of land cover from IGBP MODIS: forest, shrubland, savanna, grassland, and croplands. [@fig-studyArea]c shows the spatial distribution of the macroclasses through Chile for the year 2022. [@fig-studyArea]d shows the macroclasses of land cover persistance (80%) during 2021–2022, respectively (Table \ref{tab-landcoverSurf}). Within continental Chile, barren land is the land cover class with the highest surface area (277,870 km2). The largest type of vegetation, with 137,085 km2, is forest. Grassland (74,247 km2), savanna (55,206 km2), shrubland (25,341 km2), and cropland (3,146 km2) are the other types (Table \ref{tab-landcoverSurf}). The macrozones with major changes for 2001–2022 were "Centro," "Sur," and "Austral," with 36%, 31%, and 34% of their surface changing the type of land cover, respectively (@fig-studyArea and Table \ref{tab-landcoverTrend}). @fig-LCprop shows the summary of the proportion of surface per land cover class and macrozone, derived from the persistance mask over continental Chile.

```{=latex}
\begin{table}[!ht]
\caption{The value of Sen's slope trend next to the time-series plot of surface per land cover class (IGBP MCD12Q1.016) for 2001–2022 through Central Chile. Values of zero indicate that there was not a significant trend. Red dots on the plots indicate the maximum and minimum values of surface.}
\label{tab-landcoverTrend}
\includegraphics[]{../output/figs/table_var_landcover_macro.png}
\end{table}
```

The "Norte Chico" shows an increase in barren land of 111 $km^2 yr^{-1}$  and a reduction in the class savanna of 70 $km^2 yr^{-1}$. In the "Centro" and "Sur," there are changes with an important reduction in savanna (136 to 318 $km^2 yr^{-1}$ ), and an increase in shrubland and grassland. Showing a change for more dense vegetation types. It appears to be a shift in the area cultivated (croplands) from the "Centro" to the "Sur." Also, there is a high increase in forest (397 $km^2 yr^{-1}$ ) in the "Sur," replacing the savanna lost (Table \ref{tab-landcoverTrend}).

### Relationship between drought indices and land cover change

According to Table \ref{tab-landcoverTrendRF}, the trends in drought indices reach an r-squared between 0.32 and 0.39 for the types of forest, grassland, savanna, shrubland, and barren land. It is more likely that short- and medium-term increases in AED (EDDI-6 and EDDI-12) and short-term precipitation deficits (SPI-6 and SPEI-6) contributed to changes in grassland and bare land. The short-term increase of AED (EDDI-3 and EDDI-6) and the longer duration of the precipitation deficit (SPI-24, SPI-36, and SPEI-36) most likely contribute to the changes in shrubland. The changes in savanna are associated with a short- and long-term increase in AED and a three-year precipitation deficit (SPI-36). The increase in cumulative AED from 12 to 36 months is the strongest associated variable that contributes to changes in forests, followed by the reduction of soil moisture over six and 36 months. More details about the results of the variable importance can be found in the supplementary material in Section S3.

```{=latex}
\begin{table}[!ht]
\caption{The five most important trends of drought indices in estimating the landcover trend per land cover type and the r-squared (rsq) reached by each random forest model.}
\label{tab-landcoverTrendRF}
\includegraphics[]{../output/figs/table_importance_trends_landcover_vs_drought.png}
\end{table}
```

```{r fig-TrendsLandDrought}
#| echo: false
#| fig-pos: "!ht"
#| fig-cap: "Relationship between the trend in land cover change (y-axis) and the trend in drought indices (x-axis) for the five macrozones. Vertical panels correspond to 1, 3, 6, 12, 24, and 36 months of the time scale by drought index. Horizontal panels show each drought index"
#| out-width: "\\textwidth"
knitr::include_graphics('../output/figs/points_landcover_drought_indices_trend_and_time_scale.png')
```

We study the connection between the SPI, EDDI, and SSI drought indices and changes in land cover in @fig-TrendsLandDrought. To do this, we compare the relative changes in land cover (in terms of the total surface area per land cover type and macrozone) over six and thirty-six months. @fig-TrendsLandDrought shows that the forest in the "Sur," shrubland and grassland in "Centro," barren land in "Norte Chico," and savanna in "Austral" showed an increase in the surface of landcover associated with an increase in EDDI. Savanna in "Centro," "Sur," and "Norte Chico" decreases with the increase in EDDI. The SPI and SSI showed similar behavior regarding the trend in land cover type. A decrease in SPI and SSI is associated with an increase in the surface in shurubland and grassland in "Centro," forest in "Sur," and barren land in "Norte Chico," as well as a decrease trend in savanna in "Norte Chico," "Centro," and "Sur."

## Drought impacts on vegetation productivity within land cover

### Vegetation productivity

```{r fig-zcNDVI_var}
#| echo: false
#| fig-pos: "!ht"
#| fig-cap: "(a) Map of the linear trend of the index zcNDVI-6 for 2001–2023. Greener colors indicate a positive trend; reder colors correspond to a negative trend and a decrease in vegetation productivity. Grey colors indicate either no vegetation or a change in land cover type for 2001–2022. (b) Temporal variation of zcNDVI-6 aggregated at macrozone level within continental Chile. Each horizontal panel corresponds to a macrozone from 'Norte Grande' to 'Austral'."
#| out-width: "\\textwidth"
knitr::include_graphics("../output/figs/temporal_variation_zcNDVI6_macrozonas_con_mapa.png")
```

In @fig-zcNDVI_var it is showed the spatial map of trends in zcNDVI ([@fig-zcNDVI_var]a) and the temporal variation of zcNDVI within the aggregated macrozones ([@fig-zcNDVI_var]b). In "Norte Grande," vegetation productivity, as per the z-index, exhibits a yearly increase of 0.02 with respect to grassland and shrubland categories. There is a negative trend in "Norte Chico" with -0.04 and "Centro" with -0.02 per decade. In the "Norte Chico," savanna (-0.05) has the lowest trend, and the rest of the types are around -0.04. In "Centro," shrubland reaches -0.06, grassland -0.05, and croplands and savanna -0.01 per decade. This could be associated either with a reduction in vegetation surface, a decrease in biomass, or browning [@Miranda2023]. Vegetation reached its lowest values since the year 2019, with an extreme condition in early 2020 and 2022 in the "Norte Chico" and "Centro". The "Sur" and "Austral" show a positive trend of around 0.016 per decade (@fig-zcNDVI_var). Despite the croplands suffering from drought just as badly as the native vegetation in "Norte Chico," the savanna and shurbland appears to be the region most affected by a negative trend in vegetation productivity.

### Correlation between vegetation productivity and drought indices

```{r fig-corTimeScale}
#| echo: false
#| fig-pos: "!ht"
#| out-width: "\\textwidth"
#| fig-cap: "Time scales per drought index that reach the maximum coefficient of determination"
knitr::include_graphics("../output/figs/mapa_cor_selec_indices_zcNDVI6.png")
```

```{r fig-corPerson}
#| echo: false
#| fig-pos: "!ht"
#| out-width: "\\textwidth"
#| fig-cap: "Pearson correlation value for the time scales and drought index that reach the maximum coefficient of determination"
knitr::include_graphics("../output/figs/mapa_cor_r_indices_zcNDVI6.png")
```

@fig-corTimeScale is a map that shows the highest coefficient of determination (r-squared, or rsq) found in the regression analysis between different drought indicators and plant productivity over time. The spatial variation of time scales reached per index is mostly for time scales above 12 months. In the case of SSI, the predominant scales are 6 and 12 months. For all indices, to the north, the time scales are higher and diminish toward the south until the south part of "Austral" increases. In @fig-corPerson, the map of Pearson correlation values is shown. The EDDI reached correlations above 0.5 between "Norte Chico" and "Sur." The correlation changes from negative to positive toward the Andes Mountains and to the sea, just as in the northern part of "Austral." The SPI and SPEI have similar results, with the higher values in "Norte Chico" and "Centro" being higher than 0.6. Following a similar spatial pattern as EDDI. The SSI showed to be the index that has a major spatial extension with a higher correlation. It has a similar correlation to SPI and SPEI for "Norte Chico" and "Sur,"  but has improvements for "Sur."

```{=latex}
\begin{table}[!ht]
\caption{Summarry per land cover macroclass and macrozone regarding the correlation between zcNDVI with the drought indices EDDI, SPI, SPEI, and SSI for time scales of 1, 3, 6, 12, 24, and 36. The numbers in each cell indicate the time scale that reached the maximum correlation for the land cover and macrozone, and the color indicates the strength of the r-squared obtained with the index and the time scale.}
\label{tab-corlandcover}
\includegraphics[]{../output/figs/tabla_r_cor_macro_indice.png}
\end{table}
```

In Table \ref{tab-corlandcover}, we aggregate per macrozone and landcover the correlation analysis presented in @fig-corTimeScale and @fig-corPerson. According to what is shown, forests seem to be the most resistant to drought. Showing that only "Centro" is slightly (rsq = 0.25) impacted by a 12-month soil moisture deficit (SSI-12). In the "Norte Chico" and to a lesser extent in the "Norte Grande," it is evident that a SSI-12 with a rsq = 0.45 and a decrease in water supply (SPI-36 and SPEI-24 with rsq = 0.28 and 0.34, respectively) have an impact on grasslands. However, this type was unaffected by soil moisture, water supply, or demand in macrozones further south. The types that show to be most affected by variation in climate conditions are shrublands, savannas, and croplands. For savannas in "Norte Chico," the SSI-12 and SPI-24 reached an rsq of 0.74 and 0.58, respectively. This value decreases to the south, but the SSI-12 is still the variable explaining more of the variation in vegetation productivity (rsq = 0.45 in "Centro" and 0.2 in "Sur"). In the case of croplands, the SPEI-12, SPI-36, and SSI-12 explain between 45% and 66% of the variability in "Norte Chico." The type of land most impacted by climatic variation was shrubland, where soil moisture explained 59% and precipitation, 37%, in "Norte Chico" and "Centro," with SSI-12 being the most relevant variable, then SPI-36 in "Norte Chico" and SPI-24 in "Sur."

#	Discussion

## Drought trend and attribution to land cover

@Vicente-Serrano2022, in a study at the global scale of drought trends, indicate that there have not been significant trends in meteorological drought since 1950. Also, state that the increase in hidrological trend in some parts of the globe (northeast Brazil and the Mediterranean region) is related to changes in land cover and specifically to the rapidly increasing irrigated area, which consequently increases water extraction. @Kogan2020 analyzed the agricultural drought impact globally and in the main grain producer countries, finding that _"since 1980, the Earth warming has not changed the drought area or intensity."_ In our study, we considered the variation in vegetation productivity in Chile for areas without changes in land cover, to avoid misleading conclusions that could be related to the increase in water demand due to land cover change. Our results show a contrasting perspective. There has been a significant trend in the decline of vegetation productivity (zcNDVI) since 2000 for "Norte Chico" and "Centro," which has been extreme between 2020 and 2022, seemsly due to an intense hydrological drought due to the persistance of the mega drought [@Garreaud2017]. However, a rise in irrigated land doesn't seem to have an impact on this hydrological drought. Despite using the persistance mask for vegetation's trend analysis, cropland, which is the most water-demand type, showed a decrease trend in "Norte Chico" and "Centro." Also, there was an increase in barren land for both types. These changes are associated with a decrease in water demand from vegetation. Nonetheless, we used the persistant land cover to ensure that the pixel has the same class; in the case of croplands, it could happen that some areas had changed crops for others with higher water consumption and consequently increase water demand. But this effect should be minor compared to the results from land cover change at this scale of analysis.

On the other hand, for "Norte Chico" and "Centro," our results show a decrease in trends of water supply (SPI and SSI), which are higher at larger time scales, which is evidence of the hydrological drought. We say that what happened in central Chile goes against what has been found on a global scale [@Vicente-Serrano2022;@Kogan2020]. This shows that the main cause of the hydrological drought in Chile was a steady drop in water supply made worse by an increase in AED, but it seems that in zones most affected by drought, the main cause is not an increase in water demand by vegetation like irrigated crops. Finally, north-central Chile has experienced a decline in vegetation productivity across all macroclasses, which is primarily attributable to variations in water supply and soil moisture. An increase in water demand, such as an increase in the surface area of irrigated crops, could strengthen this trend.

## Land cover types and their impact by drought

We discovered that croplands, savannas, and shrubland are the most susceptible to climatic changes and are most affected by the 12-month soil moisture deficit. In a study in the Yangtze River Basin in China, @Jiang2020 analyzed the impact of drought on vegetation using the SPEI and the Enahanced Vegetation Index (EVI). They found that cropland was more sensitive to drought than grassland, showing that cropland responds strongly to short- and medium-term drought (< SPEI-6). In our case, the SPEI-12 was the one that most impacted the croplands in "Norte Chico" and "Centro." In general, most studies show that croplands are most sensitive to short-term drought (< SPI-6) [@Zambrano2016;@Potopova2015;@Dai2020;@Rhee2010]. Short-term precipitation deficits impact soil water, and thus less water is available for plant growth. However, we found that in "Norte Chico," an SPI-36 and SPEI-12 had a higher impact, which are associated with hydrological drought (long-term), and in "Centro," an SPI-12 and SPEI-12. Thus, we attribute this impact to the hydrological drought that has decreased groundwater storage [@Taucare2024], which in turn is impacted by long-term deficits, and consequently, the vegetation is more dependent on groundwater. In "Sur" and "Austral," the correlations between drought indices and vegetation productivity decrease, as do the time scales that reach the maximum r-squared. What can be explained is that, south of "Centro," predominate forest and grassland, the most resistant types. Also, drought episodes have been less frequent and intense. The drought episodes have had a lower impact on water availability for vegetation.

According to @Senf2020, severe drought conditions in Europe are a significant cause of tree mortality. However, we found that forest is the type of land cover macroclass less affected by variation in drought indices, being the most resistant land cover class to drought. Supporting this is @Fathi-Taperasht2022], who assert that Indian forests are the most drought-resistant and recover rapidly. Similarly, the work of @Wu2024, who analyzed vegetation loss and recovery in response to meteorological drought in the humid subtropical Pearl River basin in China, indicates that forests showed higher drought resistance. Using Vegetation Optical Depth (VOD), kNDVI, and EVI, @Xiao2023 test the resistance of ecosystems and find that ecosystems with more forests are better able to handle severe droughts than croplands. They attribute the difference to a deeper rooting depth of trees, a higher water storage capacity, and different water use strategies between forest and cropland [@Xiao2023].

In contrast to what we obtained, @Venegas2022, who studied Cryptocarya alba and Beilschmiedia miersii (both from the Lauraceae family) that live in sclerophyllous forests in Chile, found that the trees' overall growth had slowed down. This could mean that the natural dynamics of their forests have changed. They attributed it to the cumulative effects of the unprecedented drought (i.e., hydrological drought). Thus, we attribute that forest to being the most resistant to drought, due to the fact that most of the species comprising it are highly resilient to water scarcity compared to the other land cover classes. Nonetheless, if we want to go deep in our analysis, we should use earth observation data that is able to capture a higher level of detail. For example, when we used MOD13A3 with a 1km spatial resolution to measure vegetation condition, it took the average condition of 1 square kilometer. Then, to use remote sensing to look at how a certain type of forest (like sclerophyllous forest) changes in response to drought on a local level, we should use operational products with higher spatial resolutions, like those from Landsat or Sentinel. This will let us do a more thorough analysis.

## Soil moisture, vegetation productivity, and agricultural drought.

The main external factors that affect biomass production by vegetation are actual evapotranspiration and soil moisture, and the rate of ET in turn depends on the availability of water storage in the root zone. Thus, soil moisture plays a key role in land carbon uptake and, consequently, in the production of biomass [@Humphrey2021]. Moreover, @Zhang2022 indicate there is a bidirectional causality between soil moisture and vegetation productivity. Lastly, some studies have redefined agricultural drought as soil moisture drought from a hydrological perspective [@Loon2016;@Samaniego2018]. Even though soil moisture is the external factor most determinant of vegetation biomass, there are multiple internal factors, such as species, physiological characteristics, and plant hydraulics, that would affect vegetation productivity. Because of that, we believe that agricultural drought, referring to the drought that impacts vegetation productivity, is the most proper term, as originally defined by @Wilhite1985.

The study results showed that the soil moisture-based drought index (SSI) was better at explaining vegetation productivity across land cover macroclasses than meteorological drought indices like SPI, SPEI, and EDDI. In the early growing season and especially in irrigated rather than rainfed croplands, soil moisture has better skills than SPI and SPEI for estimating gross primary production (GPP). This according to @Chatterjee2022 evaluation of the SPI and SPEI and their correlation with GPP in the CONUS. Also, @Zhou2021 indicate that the monthly scaled Standardized Water Deficit Index (SWDI) can accurately show the effects of agricultural drought in most of China. @Nicolai2017 also looked at the time-lag between the SWDI and the Vegetation Condition Index (VCI). They found that there was little to no time-lag in croplands but a greater time-lag in forests.

In our case, there is strong spatial variability throughout Chile and between classes, mainly attributable to climate heterogeneity, hydrological status, or vegetation resistance to water scarcity. The semi-arid "Norte Chico" and the Mediterranean "Centro" were where SSI had the best performance. In Chile, medium-term deficits of 12 months are more relevant in the response of vegetation, which decreases to the south, and in the case of croplands, they seem to react in a shorter time, with six months (SSI-6) in "Centro." This variation for croplands could be related to the fact that in "Norte Chico," the majority of crops are irrigated, but to the south there is a higher proportion of rainfed agriculture, which is most dependent on the short-term availability of water. Rather, in the "Norte Chico," the orchards are more dependent on the storage of water in dams of groundwater reservoirs, which are affected by long-term drought (e.g., SPI-36).

## Future outlook (to complete)

# Conclusion

There is a trend toward decreasing water supply in most parts of Chile, particularly in the “Centro” and “Norte Chico” regions. The whole country showed an increase in AED. Vegetation productivity only showed a decrease in the “Norte Chico” and “Centro,” being highest for shrubland and croplands. Forest is the land cover most resistant to drought, as shown along Chile, and shrubland and cropland are the most sensitive.

A soil moisture deficit of 12 months (SSI-12) is highly correlated with vegetation productivity for the land cover classes of shrubland, savannas, croplands, and forest in “Norte Chico” and “Centro.” For the southern part of the country with humid conditions, the correlation with SSI decreases. Soil moisture overcomes the capacity to explain vegetation productivity over the supply and demand drought indices in the entire territory.

The variation in vegetation productivity appears to be associated with climate variation rather than anthropogenic factors (e.g., an increase in water demand by irrigated crops). Even though switching to more demanding crops on the land could increase the impact of drought on vegetation, this would need to be more thoroughly investigated, for instance at the watershed level.

The results of this study could help in the development of a robust forecasting system for land cover classes in Chile, helping to improve preparedness for climate change impacts on vegetation.

# Supplementary material {-}

# References {-}